---
title: "ACCS-UAA Water Temp Data Import"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Benjamin Meyer (benjamin.meyer.ak@gmail.com)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear environment
rm(list=ls())

# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
```


Read in data from ACCS folder
```{r}
accs.data <-list.files(path = paste0("data/ACCS-UAA/"),
               pattern = "*.csv", 
               full.names = T) %>% 
    map_df(~read_csv(., col_types = cols(.default = "c"),col_names = T)) 

# note: some files have a "location description" column and others do not

# rename, transform, and eliminate columns to match final field names described in Project_notes.Rmd
accs.data <- accs.data %>%
  select(AKOATS_ID,sampleDate,sampleTime,Temperature) %>%
  transform(sampleDate = as.Date(mdy(sampleDate)),
            sampleTime = hms::as.hms(sampleTime),
            Temperature = as.numeric(Temperature)) 


# for visualization and data QA/QC, we want waterbody name and/or SiteID associated with each dataset

# read in metadata file and associate SiteID with AKOATS_ID
md <- read.csv("data/ACCS-UAA/metadata/SiteLevelMetadata_Bogan.csv") %>%
  select(AKOATS_ID,SiteID,Waterbody_name)

# use info from pdf at "data/ACCS-UAA/metadata/doi_10.5063_F1028PR9-METADATA.pdf" to assign waterbody names in blank spaces
md.x <- md %>%
  filter(AKOATS_ID <= 764) %>%
  select(-Waterbody_name)

# 760 = Koktuli River
# 761 = Tributary of Koktuli River 1
# 762 = Tributary of Koktuli River 2
# 763 = tributary of Koktuli River 3
# 764 = tributary of Newhalen River

# create completed metadata table
Waterbody_name <- c("Koktuli River", "Tributary of Koktuli River 1", "Tributary of Koktuli River 2","Tributary of Koktuli River 3","Tributary of Newhalen River")
AKOATS_ID <- 760:764
md.x1 <- data.frame(Waterbody_name,AKOATS_ID)

md.x <- left_join(md.x,md.x1,by = "AKOATS_ID")

# rejoin metadata w/ manually input site names back to UAA-ACCS metadata
md <- md %>%
  filter(AKOATS_ID > 764) %>%
  bind_rows(md.x) %>%
  transform(AKOATS_ID = as.character(AKOATS_ID))

# associate SiteID and Waterbody_name with overall UAA-ACCS temperature dataset
accs.data <- left_join(accs.data,md,by = "AKOATS_ID")
  
```

<br>

Perform a quick visualization to see extent and form of data
```{r}
accs.data %>%
  select(-sampleTime) %>%
  mutate(year = as.factor(year(sampleDate)),
         day = yday(sampleDate)) %>%
  group_by(day,Waterbody_name,SiteID,year) %>%
  summarise(daily_mean_Temperature = mean(Temperature)) %>%
  ggplot(aes(day,daily_mean_Temperature, color = year)) +
  geom_point() +
  facet_wrap(. ~ Waterbody_name) +
  ggtitle("Original Logger Data by Site and Year")

```

<br>

Looks like we have some malfunctioning loggers at site Bonanza Creek.  Lets identify 2015 and 2016 data from this site as not usable by identifying with a "0" in the "useData" column.

```{r}
# FYI this can be done in a more fine-tuned fashion; eliminating specific data by day or individual observation(s), by creating this file externally as a csv and listing individual  observations to be eliminated
year <- c(2015, 2016)
AKOATS_ID <- c("1664","1664")
useData <- c(0,0)
bad_data <- data.frame(year,AKOATS_ID,useData) %>%
  transform(useData = as.character(useData))

accs.data <- accs.data %>%
  left_join(bad_data) %>%
  mutate(useData = replace_na(useData, "1")) %>%
  distinct() %>%
  mutate(year = year(sampleDate))

```

<br>

Perform a quick visualization of "cleaned up" data
```{r}
accs.data %>%
  filter(useData == 1) %>%
  select(-sampleTime) %>%
  mutate(year = as.factor(year(sampleDate)),
         day = yday(sampleDate)) %>%
  group_by(day,Waterbody_name,SiteID,year) %>%
  summarise(daily_mean_Temperature = mean(Temperature)) %>%
  ggplot(aes(day,daily_mean_Temperature, color = year)) +
  geom_point() +
  facet_wrap(. ~ Waterbody_name) +
  ggtitle("Corrected Logger Data by Site and Year")

```

<br>

Improvement is evident, probably more cleaning to be done at sub-daily mean level.

<br>

Interactive data summary table
```{r}
# create data summary table
accs.data.summary <- accs.data %>%
  filter(useData == 1) %>%
  group_by(SiteID,AKOATS_ID,Waterbody_name,year) %>%
  summarize(meanTemp = mean(Temperature, na.rm = T),
            maxTemp = max(Temperature, na.rm = T),
            minTemp = min(Temperature, na.rm = T),
            sdTemp = sd(Temperature, na.rm = T),
            n_obs = n())

accs.data.summary %>%
  datatable() %>%
  formatRound(columns=c("meanTemp","maxTemp","minTemp","sdTemp"), digits=2)
    
```

<br>

Export csv of combined datasets, with newly identified "do not use data" signified with a "0" in the useData column, to data folder associated with this .Rproj
```{r}
# reorder columns
x <- accs.data %>% select(AKOATS_ID,SiteID,sampleDate,sampleTime,Temperature,useData)

# export csv
write.csv(x,"output/UAA-ACCS.csv")

```
