---
title: "NPS Water Temp Data"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Benjamin Meyer (benjamin.meyer.ak@gmail.com)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear environment
rm(list=ls())

# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
```

<br>

Read in data from folder "NPS Bartz"
```{r}
# create dataframe of multi-depth-loggers Lake sites to read in later
z <- read.csv(
)

# site-files that we want to read in seperately are assigned "Sensor_placement = Lake Center" in the Site Info fale; use this to filter

nps.data <-list.files(path = paste0("data/NPS Bartz"),
               pattern = "*.csv", 
               full.names = T) %>%
  data.frame() %>%
  filter()
nps.data <- nps.data[!grepl(c("Pilot"), nps.data)] 
z <- nps.data %>%
  map_df(~read_csv(., col_types = cols(.default = "c"),col_names = T, skip = 15)) 

```

Notes: 

* The file "All_Pilot_Sites_20201104.csv" is not in a format consistent with all other single-logger-per-site csv files.  Read in separately.

* The following csv files are site files containing all data from an anchored line in a lake w/ loggers at various (5m-10m) depth increments.  Read in and address these separately:

- LCLAR_01_TEMP
- KIJIL_01_TEMP
- TELAL_01_TEMP
- LBROO_01_TEMP
- NAKNL_03_TEMP
- NAKNL_01_TEMP
- NAKNL_02_TEMP


* Among sites and years, various types of loggers are employed (sonde, level logger, HOBO, etc)


