---
title: "data_Deshka_Temperature_Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggrepel)
library(lubridate)
library(readxl)
library(rgdal)
# library(sp)
# library(sf)
library(raster)
library(tidyverse)

capwords <- function(s, strict = FALSE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                  {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

```

This report is for reading in the data used in the Deshka stream temperature model. I want to read in all data as originally provided and combine it into one file for import into AKTEMP. I am copying over the code used to merge the data files together. No data were flagged for this project. (Note: for the model, some sites or years were removed, but this is the complete dataset as originally provided.)

Project funded by Mat-Su Salmon Habitat Partnership, award no. F18AC00942
Ended 6/30/20.

Data are from CIK (~ 60 sites, 2017-2019) and FWS (~30 sites, 2019 only at minnow-trapping locations).

```{r}
deshka_folder <- "W:/Github/Deshka_temperature/"
```

# Cook Inletkeeper Data

## Read and combine excel files

Read in all excel files in the data folder. Sue included some files that show all available data (metadata summaries); remove those from vector of file names before reading in data.

```{r}

files <- list.files(path = paste(deshka_folder, "data/", sep = ""), pattern = ".xlsx", full.names = TRUE)

files <- files[!files %in% (grep("availability", files, value = TRUE))]

```

Sheet names are the site names, although it looks like some years are broken into separate sheets.
Create a function that reads in each sheet and uses the sheet name to assign the site_ID.

```{r}
combine_sheets_with_site_ID <- function(excel_file_path) {
  sheets <- excel_sheets(excel_file_path)
  all_dat <- tibble()
  for(i in sheets){
    dat <- read_excel(excel_file_path, sheet = i) %>% 
      mutate(SiteID = i)
    all_dat <- bind_rows(all_dat, dat)
  }
  return(all_dat)
}
```

Read in all of the excel files into a data frame using map.

```{r}
temp <- files %>% 
  map_df(function(x) combine_sheets_with_site_ID(x)) 
```

Rename fields and modify date fields.

```{r}
temp <- temp %>% 
  mutate(sampleDate = as.Date(Date, format = "%Y-%m-%d"),
         sampleTime = strftime(`Time, GMT-08:00`, format = "%H:%M"),
         useData = 1) %>% 
  rename(Temperature = `Temp, °C`) %>% 
  filter(!is.na(Temperature)) %>% 
  dplyr::select(SiteID, sampleDate, sampleTime, Temperature, useData)
```

Fix site names: remove extra dates, remove extra white space, correct different tributary spellings and capitalization, make all capwords. Total of 61 sites, matches locations provided by Sue.

```{r}

site_ID_fix <- temp %>% 
  distinct(SiteID) %>% 
  mutate(site_ID2 = sub('^([^ ]+ [^ ]+ [^ ]+).*', '\\1', SiteID), #remove extra date info from Sue's sheet names
         site_ID2 = trimws(site_ID2),
         site_ID2 = capwords(site_ID2),
         site_ID2 = sub("\\bTrib\\b", "Tributary", site_ID2),
         site_ID2 = sub("TOP", "Top of Study", site_ID2))
         
site_ID_fix %>% arrange(site_ID2)
site_ID_fix %>% distinct(site_ID2)
         
temp <- temp %>% 
  left_join(site_ID_fix) %>% 
  dplyr::select(-SiteID) %>% 
  rename(SiteID = site_ID2) 

temp %>% distinct(SiteID)
```




# USFWS Road Sites

These are road-accessible sites that were logged June-August 2019 while out collecting fish data.

## Read and combine excel files

Get list of files.

```{r}
files <- list.files(path = paste(deshka_folder, "data/Road_sites/XLSX Clean Files", sep = ""), 
                    pattern = ".xlsx", full.names = TRUE)
```

Read in data.

```{r}
rdtemp <- files %>% 
  map_df(function(x) {
    dat <- read_excel(path = x, skip = 2, col_names = c("Date", "Time", "Temperature"), col_types = c("date", "date", "numeric"))
    name <- str_extract(excel_sheets(x), "[^_]+")
    newdat <- dat %>% 
      mutate(SiteID = name)
    return(newdat)
  }
  ) 

rdtemp <- rdtemp %>% 
  mutate(sampleDate = as.Date(Date, format = "%Y-%m-%d"),
         sampleTime = strftime(Time, format = "%H:%M"),
         SiteID = case_when(SiteID == "OWl1" ~ "OWL1", 
                             TRUE ~ SiteID),
         useData = 1) %>% 
  select(SiteID, sampleDate, sampleTime, Temperature, useData)

rdtemp %>% distinct(SiteID)

```


# Combine datasets and save


```{r}

names(temp)
names(rdtemp)

bind_rows(temp, rdtemp) %>% 
  write_csv(path = "output/deshka_90sites.csv")

```



