---
title: "Cook Inletkeeper Water Temps"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Rebecca Shaftel (rsshaftel@alaska.edu).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))



# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
library(dataRetrieval)
```

# Metadata

AKOATS metadata

```{r}
akoats.meta <- read_excel("data/AKOATS_DATA_2020_Working.xlsx", sheet = "CONTINUOUS_DATA", col_types = "text") %>%
  select(seq_id,Agency_ID,Contact_person,SourceName,Contact_email,Contact_telephone,Latitude,Longitude,Sensor_accuracy,Waterbody_name) %>%
  rename(AKOATS_ID = seq_id,
         SiteID = Agency_ID) %>% 
  mutate(AKOATS_ID = as.numeric(AKOATS_ID),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

akoats.meta

```

CIK metadata

Sue didn't provide lat/longs in her metadata file so need to try and match names in akoats. Sue provided lat/long for Ben Courtney Creek, Dan provided lat/long for Roadhouse creek.

```{r}

cik.meta <- akoats.meta %>%
  filter(SourceName == "CIK") %>% 
  select(SiteID, SourceName, Contact_person, Waterbody_name, Latitude, Longitude)

#remove sites in cook inlet.
cik.meta <- cik.meta[!grepl("CIK", cik.meta$SiteID),]

cik.meta <- bind_rows(cik.meta,
                      data.frame(SiteID = "Ben Courtney Creek",
                                 SourceName = "CIK",
                                 Contact_person = "Sue Mauger",
                                 Latitude = 59.272417,
                                 Longitude = -156.357567,
                                 Waterbody_name = "Ben Courtney Creek"))

cik.meta <- bind_rows(cik.meta,
                      data.frame(SiteID = "Roadhouse Creek",
                                 SourceName = "CIK",
                                 Contact_person = "Sue Mauger",
                                 Latitude = 59.7573,
                                 Longitude = -154.8499,
                                 Waterbody_name = "Roadhouse Creek"))

cik.meta <- cik.meta %>% 
  mutate(SiteID = case_when(grepl("Gibraltar", SiteID) ~ "Gibraltar River",
                            TRUE ~ SiteID))
  
saveRDS(cik.meta, "output/cik_metadata.rds")

```



# Data

Read in temp observation data from CIK folder

```{r}
cik.files <-list.files(path = paste0("data/CIK Bristol Bay village and lodge sites"),
               pattern = "*.xlsx", 
               full.names = T)

#remove metadata file
cik.files <- cik.files[!(grepl("CIK Bristol Bay village and lodge sites/CIK Bristol", cik.files))]


cik.data <- cik.files %>% 
  map_df(function(x) read_excel(x,skip = 1,col_names = F) %>% 
  mutate(file_name = gsub(".xlsx","",basename(x))))

cik.data <- cik.data %>% 
  mutate(sampleDate = as.Date(`...1`),
         sampleTime = as_hms(`...2`),
         Temperature = `...3`,
         SiteID = sub("\\_.*", "", file_name)) %>%  
  select(-`...1`, -`...2`, -`...3`, -`...4`, -`...5`) 

#fix site names. 
cik.data <- cik.data %>% 
  mutate(SiteID = case_when(grepl("Naknek", file_name) ~ "Big Creek_Naknek",
                            grepl("Roadhouse", file_name) ~ "Roadhouse Creek",
                            grepl("Panar", file_name) ~ "Panaruqak Creek",
                            TRUE ~ SiteID))


```

Summary csv to add as attributes to leaflet map.

```{r}
cik.data %>% 
  group_by(SiteID) %>% 
  summarize(startYear = min(year(sampleDate)),
            endYear = max(year(sampleDate)),
            totYears = length(unique(year(sampleDate)))) %>% 
  saveRDS("output/cik_data_summ.rds")

```

```{r}
cik.data %>% 
  saveRDS("output/cik_data.rds")
```

Perform a quick visualization to see extent and form of data
```{r eval = FALSE}
cik.data %>%
  mutate(DT = parse_date_time(paste(sampleDate, sampleTime), "Y-m-d H:M:S")) %>%
  ggplot(aes(x = DT, y = Temperature)) +
  geom_point() +
  facet_wrap(~ SiteID) +
  ggtitle("Original Logger Data by Site and Year")

cik.data %>% distinct(SiteID)

```

