---
title: "data_NPS_DanYoung"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data. They also provided USGS data for the Chulitna River, but I have that already in the nwis download.

```{r}
dy.files <- list.files("S:/Stream Temperature Data/NPS Dan Young/Dan_Young_Data", pattern = ".csv", full.names = TRUE) 

#one site has different formatting.
dy.files <- dy.files[!grepl("LACL", dy.files)]

dy.dat <- map_df(dy.files, function(x) read_csv(x, comment = "#") %>% mutate(file_name = basename(x)))

dy.dat %>% distinct(Grade)
dy.dat %>% distinct(Qualifiers)
dy.dat %>% distinct(`Approval Level`)

dy.dat <- dy.dat %>% 
  mutate(sampleDate = as.Date(`Timestamp (UTC-08:00)`),
         sampleTime = as_hms(`Timestamp (UTC-08:00)`),
         dt = as.POSIXct(paste(sampleDate, sampleTime), format = "%Y-%m-%d %H:%M:%S"),
         SiteID = gsub(".csv","",basename(file_name)),
         UseData = 1) %>%
  mutate(SiteID = case_when(SiteID == "Chulitna_Dan_2007_2011" ~ "Chulitna_SiteB",
                            TRUE ~ SiteID)) %>% 
  rename(Temperature = Value) %>% 
  select(file_name, SiteID, sampleDate, sampleTime, dt, Temperature, UseData)

```

This is a site on the Chulitna River that corresponds to an old USGS gage location.

```{r}
dy.lacl <- read_csv("S:/Stream Temperature Data/NPS Dan Young/Dan_Young_Data/LACL_chulr_temp_20200925export.csv", skip = 1)

dy.lacl <- dy.lacl %>% 
  rename(sampleTime = `Time, GMT-08:00`,
         Temperature = `Temp, °C (LGR S/N: 11001174, SEN S/N: 11001174, LBL: LACL_CHULR_TEMP_)`) %>% 
  mutate(sampleDate = as.Date(Date, format = "%m/%d/%Y"), 
         SiteID = 15298040,
         dt = as.POSIXct(paste(sampleDate, sampleTime), format = "%Y-%m-%d %H:%M:%S"),
         UseData = 1) %>% 
  select(SiteID, sampleDate, sampleTime, Temperature, UseData)
```

Look at the data:
* Chulitna R looks like it is getting dewatered very regularly. I would throw this site out altogether except anything that might be useful from 2018 could be combined with the historic USGS data. 2019 and 2020 are out -- too much bad data.


```{r}
p1 <- dy.lacl %>% 
  ggplot(aes(x = dt, y = Temperature)) + 
  geom_line()

dy.sites <- dy.dat %>% distinct(SiteID) %>% pull(SiteID)

p2 <- dy.dat %>% 
  filter(SiteID == dy.sites[2]) %>% 
  ggplot(aes(x = dt, y = Temperature)) + 
  geom_line()


ggplotly(p2)
```

