---
title: "ACCS-UAA Water Temp Data Import"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---
Document last updated `r Sys.time()` by Rebecca Shaftel (rsshaftel@alaska.edu).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))

# source(file = "Temp_flags_function.R")


# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
```

# Metadata

Copied over Dustin's md table from the Bristol Bay Temperature repo. Create this data frame first because it has all 15 sites in the clean data file along with the akoats ids. Then link to the akoats metadata to get lat/longs.

```{r AKOATS, table}
# Create akoats metadata table using new site ids - Big Creek:ilbig01,

accs_md <- tibble(
  AKOATS_ID = as.numeric(c("1660", "1661", "1662", "1663",
                "1664", "1665", "1666", "1667",
                "1668", "760", "761", "762", 
                "763", "764", "1784")),
  SiteID = c("AKBB-020", "AKBB-028", "AKBB-029", 
             "AKBB-040", "mubon10", "musfk01",
             "mustu17", "iltaz02", "AKBB-011",
             "mussm15", "mutsk09", "mutsk02", 
             "muekm23", "iltnr19","ilbig01"),
  Waterbody_name = c("Little Mulchatna River",
                     "Unnamed Tributary Koksetna River", "Kaskanak Creek",
                     "Chilchitna River", "Bonanza Creek",
                     "South Fork Koktuli River", "Stuyahok River",
                     "Tazimina River", "Victoria Creek", "Koktuli River",
                     "Tributary of Koktuli River 1",
                     "Tributary of Koktuli River 2",
                     "Tributary of Koktuli River 3",
                     "Tributary of Newhalen River", "Big Creek")
)

accs_md%>% 
  arrange(desc(SiteID))

```


```{r}

akoats_md <- read_excel("data/AKOATS_DATA_2020_Working.xlsx", sheet = "CONTINUOUS_DATA") %>%
  select(seq_id,Agency_ID,Contact_person,SourceName,Contact_email,Contact_telephone,Latitude,Longitude,Sensor_accuracy) %>%
  rename(AKOATS_ID = seq_id,
         SiteID = Agency_ID) %>% 
  mutate(AKOATS_ID = as.numeric(AKOATS_ID),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

#Site IDs changed so just join on akoats id.
accs_md <- left_join(accs_md, akoats_md %>% select(-SiteID))

saveRDS(accs_md, "output/accs_metadata.rds")
```


# Data

Read in Dustin's clean data file from the Bristol Bay Temperature repo.

```{r}

accs.data <- read_csv("W:/Github/Bristol-Bay-Temperature/output/BB_Temp_Clean_2020-12-18.csv")

accs.data %>% 
  distinct(SiteID, sampleDate) %>% 
  count(SiteID, year = year(sampleDate)) %>% 
  arrange(year) %>% 
  pivot_wider(names_from = year, values_from = n)
```

For visualization and data QA/QC, we want waterbody name associated with each dataset, which we need from the metadata. 

```{r}

accs.data <- left_join(accs.data, accs_md %>% select(SiteID, Waterbody_name), by = "SiteID")

accs.data %>% distinct(Waterbody_name, SiteID)


```

Summary csv to add as attributes to leaflet map.

```{r}
accs.data %>% 
  group_by(SiteID, Waterbody_name) %>% 
  summarize(startYear = min(year(sampleDate)),
            endYear = max(year(sampleDate)),
            totYears = length(unique(year(sampleDate)))) %>% 
  saveRDS("output/accs_data_summ.rds")

```


Save a copy of the data for a summary table and figure.

```{r}
saveRDS(accs.data, "output/accs_data.rds")
```




# Data review

Perform a quick visualization to see extent and form of data. Note that these data have now been qaed by Dustin.

```{r}

accs.data %>%
  filter(UseData == 1) %>% 
  mutate(DT = parse_date_time(paste(sampleDate, sampleTime), "Y-m-d H:M:S")) %>%
  ggplot(aes(x = DT, y = Temperature)) +
  geom_point() +
  facet_wrap(~ SiteID) +
  ggtitle("Original Logger Data by Site and Year")

```

