---
title: "1_data_exploration"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 10
    fig_caption: yes
    code_folding: hide
    toc: true
    number_sections: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
date: "2023-01-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(sf)
library(lubridate)
library(broom)
library(cowplot)
library(ggrepel)
library(zoo)
library(trend)
library(Kendall)
library(ggfortify)
```

# Read in data and add attributes

All data were combined in the AKSSF repository. Read from that project drive and filter on data from Bristol Bay.

NOTE: Data from Zenodo are incomplete and are missing the UW data from 2017 on.

```{r zenodo data, eval = FALSE}
temp <- read_csv(file =  "C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\zenodo1\\dailyTemps.csv")

mets <- read_csv(file = "C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\zenodo1\\tempMetrics.csv")

md <- read_csv(file = "C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\zenodo1\\tempSitesMetadata.csv")

names(md)
```

Try a different dataset that Tim used for the DFA -- all years are there!

```{r DFA input data}

temp <- read_csv("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\summer_data_wair_dayl2022-03-17.csv")
md <- read_csv("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\md_2022-02-08.csv")

md %>% count(Waterbody_type)
md %>% filter(Region == "Bristol Bay") %>% distinct(SourceName, Contact_person)
```

Adding watershed attributes created for the AKSSF project to these sites.

```{r read in TS model data frame}
mod_dat <- readRDS("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\model_data2022-05-09.rds")

mod_dat

#note that Tim had me populate all airtemps for this data frame, can remove dates without
# meandt
bb_temp <- left_join(temp, md %>% 
                       select(Site, SourceName, Waterbody_name, Waterbody_type, Region, Name, Latitude, Longitude, Waterbody_name, HUC8), by = c("SiteID" = "Site")) %>% 
    filter(Region == "Bristol Bay", !is.na(meanDT)) 

#9 sites in temperature data frame that we didn't calculate geospatial data for.
# I think these were dropped bc too little data or incomplete summer data.
anti_join(bb_temp %>% distinct(SiteID), mod_dat %>% distinct(Site), by = c("SiteID" = "Site"))

names(mod_dat)

bb_temp <- left_join(bb_temp, mod_dat %>% 
            select(Site, str_ord:wtd_area_sqKM) %>% 
            distinct(), 
          by = c("SiteID" = "Site")) %>% 
  filter(!is.na(str_ord))


```


```{r BB data summary}

#creating site/date/mean daily temp frame for Erik and Will
mean_dailies<-bb_temp%>%
  select(SiteID, Latitude, Longitude, HUC8, Waterbody_name, sampleDate, meanDT)

write.csv(mean_dailies, file="output/BBDailyTemps.csv")

bb_temp %>% 
  filter(Region == "Bristol Bay", grepl("UW", SiteID)) %>% 
  distinct(SiteID, year = year(sampleDate)) %>% 
  group_by(SiteID) %>% 
  summarize(years = toString(year))

#interesting, I thought the TR/TS dataset was filtered to sites with 3 or more years of data
bb_temp %>% 
  filter(Region == "Bristol Bay") %>% 
  distinct(SiteID, year = year(sampleDate)) %>% 
  count(SiteID) %>% 
  arrange(n) 


```

Summary of sites that Daniel has coho salmon size data for.

Whitefish
Yako
Bear
Silver Salmon
Fifer
Lynx (main stream and below lake trib, which are two different temperature sites).
Hidden (maybe not enough fish...)


```{r plot of ST for sites with coho data}

coho_sites <- c("UW_Aleknagik Pfifer Creek", "UW_Aleknagik Silver Salmon Creek", 
                "UW_Aleknagik Yako Creek",
                "UW_Aleknagik Big Whitefish Creek", "UW_Nerka Bear Creek", 
                "UW_Nerka Hidden Lake Creek",
                "UW_Nerka Lynx Creek", "UW_Nerka Lynx Creek Cold Tributary", 
                "UW_Nerka Lynx Lake Tributary",
                "UW_Aleknagik Bear Creek")

bb_temp %>% 
  filter(SiteID %in% coho_sites) %>%
  complete(SiteID, sampleDate) %>% 
  mutate(doy = format(sampleDate, "%j"),
         year = year(sampleDate)) %>%  
  ggplot() +
  geom_line(aes(x = as.Date(doy, format = "%j"), y = meanDT, color = SiteID), size = 0.5) +
  facet_wrap(~year) +
  geom_hline(aes(yintercept = 18), color = "red") + 
  theme_bw() +
  labs(x = "", y = "Mean Daily Temperature (C)") +
  theme(legend.position = "bottom") +
  guides(color=guide_legend(nrow=5,byrow=TRUE))

ggsave("output/coho_sites_daily_temps.jpeg")
```


# Site summaries

## Map

Map of sites for each HUC8 in Bristol Bay. (Looks like the metadata I created already has HUC8 on there.)

```{r BB sites by huc8}

md_sf <- st_as_sf(md, coords = c("Longitude", "Latitude"), crs = "WGS84")

huc8 <- st_read(dsn = "S:\\GIS_Layers\\GIS_data\\inlandWaters\\Hydrography\\NHD\\WBD\\WBD_19_GDB.gdb", layer = "WBDHU8")
# # st_crs(huc8) 
huc8_wgs84 <- st_transform(huc8, crs = "WGS84")
# 
# # st_crs(md_sf) == st_crs 
# 
# md_sf <- st_join(md_sf, huc8_wgs84)


bb_huc8_names <- md_sf %>% 
  filter(Region == "Bristol Bay") %>% 
  st_drop_geometry() %>% 
  count(Name) %>% 
  mutate(name_lab = paste0(gsub("River", "R.", Name), " (", n, ")"))



ggplot() +
  geom_sf(data = huc8_wgs84 %>% right_join(bb_huc8_names), aes(fill = name_lab)) +
  geom_sf(data = md_sf %>% filter(Region == "Bristol Bay")) +
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(fill = "HUC8 Name", title = "Bristol Bay Sampling Sites")

ggsave("output/SWIM_figures/map of BB sites.jpeg", width = 9, height = 4, units = "in")
```


## Sites by watershed attributes and years


What sites have been monitored or are being monitored?

- number of sites in each HUC8 (possibly filter to exclude those with < 3 years of data): sites by huc8 code chunk
- number of years of data for each site: site count by years of data code chunk

Do sites represent the different stream sizes/types in Bristol Bay?

- sites by stream order or elevation.
- compare this histogram to Dustin's summary of stream order or elevation for all HUC12 outlets: sites by stream order code chunk

Do sites represent different thermal regimes?

- look at BB sites in pca biplot colored by different TR: pca biplot of thermal regimes code chunk

```{r site count by years of data}
bb_temp %>% 
  filter(month(sampleDate) %in% 6:8) %>% 
  group_by(SiteID, year) %>% 
  mutate(yr_ct = n()) %>% 
  filter(yr_ct > 73) %>% 
  distinct(SiteID, year) %>% 
  ungroup() %>% 
  count(year) %>%
  # arrange(desc(n))
  ggplot(aes(x = year, y = n)) +
  geom_col() +
  theme_bw() +
  theme() +
  labs(x = "Year", y = "Count of Sites")

ggsave("output/SWIM_figures/site count by year.jpeg", width = 4.5, height = 4.5, units = "in")
```


```{r sites by stream order}
bb_temp %>% 
  filter(str_ord > 5) %>% 
  distinct(SiteID, str_ord) %>% 
  filter(!is.na(str_ord)) %>% 
  count(str_ord)

#read in Dustin's data frame with stream order for all huc12 outlets
# bb_huc12 <- read_csv("data/spatial/AKSSF_awcHuc12s_covariates.csv")
bb_huc12 <- read_csv("data/spatial/AKSSF_AwcHuc12s_strwtd_cv.csv") %>% 
  filter(region == "Bristol_Bay")
names(bb_huc12)  

bb_huc12 %>%
  count(str_ord)

bind_rows(bb_huc12 %>% select(SiteID = cat_ID_con, str_ord) %>% mutate(type = "HUC12"),
          bb_temp %>% select(SiteID, str_ord) %>% mutate(type = "Monitoring Sites")) %>% 
  ggplot() +
  geom_histogram(aes(x = (str_ord), 
                     y=stat(density*width), fill = type), 
                 position = position_dodge()) 

bind_rows(bb_huc12 %>% select(SiteID = cat_ID_con, str_ord) %>% mutate(type = "HUC12 Outlets"),
          bb_temp %>% select(SiteID, str_ord) %>% mutate(type = "Monitoring Sites")) %>%
  count(type, order = factor(str_ord)) %>% 
  group_by(type) %>% 
  mutate(n = prop.table(n)) %>% 
  ggplot(aes(x = order, y = n, fill = type)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Stream Order", y = "Percent of Sites", fill = "Site Type")

ggsave("output/SWIM_figures/site percent by stream order.jpeg", width = 4.5, height = 4.5, units = "in")

```

```{r}

ggplot() +  
  geom_histogram(aes(x = as.factor(str_ord), y=..count../sum(..count..)), stat = "count", data = bb_huc12) +
  geom_histogram(aes(x = as.factor(str_ord + 0.2), y=..count../sum(..count..)), 
                 data = bb_temp %>% distinct(SiteID, str_ord), fill = "blue", stat = "count") +
  scale_x_discrete(breaks = c(2, 3, 4, 5, 6, 7, 8, 9))
```

```{r large rivers}
names(bb_temp)

bb_temp %>% 
  distinct(SiteID, Name, year, Waterbody_name, str_ord) %>%
  group_by(SiteID, Name, Waterbody_name, str_ord) %>% 
  summarize(years = toString(year)) %>% 
  filter(str_ord > 6) %>% 
  arrange(Waterbody_name) %>%
  ungroup() %>% 
  select(Waterbody_name, str_ord, years)
```



# Time series figures

Plots of daily temps by HUC8 and with 18C threshold for thermal stress to adults and juveniles.

```{r pdf of daily temps by huc8 with 18C threshold}

# pdf("output/Daily summer temperatures by HUC8.pdf")

for(i in bb_huc8_names) {
  dat <- left_join(temp, md %>% select(Site, Region, Name), by = c("SiteID" = "Site")) %>% 
    filter(Region == "Bristol Bay", Name == i) %>%
    complete(SiteID, sampleDate) %>% 
    mutate(doy = format(sampleDate, "%j"),
           year = year(sampleDate))
  p1 <- dat %>% 
    ggplot() +
    geom_line(aes(x = as.Date(doy, format = "%j"), y = meanDT, group = SiteID), color = "dark gray", size = 0.5) +
    facet_wrap(~year) +
    geom_hline(aes(yintercept = 18), color = "red") + 
    theme_bw() +
    labs(x = "", title = i)
  print(p1)  
}
# dev.off()


```


Single point in September at kijilo causing line to connect.

```{r}
bb_temp %>% 
  mutate(mo_day = format(sampleDate, "%m-%d"),
         site_year = paste0(SiteID, year),
         is_19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0)) %>% 
  filter(month(sampleDate) %in% 6:9) %>%
  filter(Name == "Lake Clark", year == 2019) %>% 
  complete(SiteID, sampleDate) %>%
  filter(grepl("kijilo", SiteID)) #%>% 
  ggplot() +
  geom_line(data = . %>% filter(is_19 == 0), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year), color = "grey") +
  geom_line(data = . %>% filter(is_19 == 1), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, color = site_year)) +
  geom_abline(aes(intercept = 18, slope = 0), linetype = 2) +
  facet_wrap(~Name) +
  theme_minimal() +
  theme(legend.position = "right", axis.title.x = element_blank(), text = element_text(size = 16)) +
  labs(y = "Mean Daily Temperature (°C)")
```


```{r time series by huc8}

b1 <- bb_temp %>% 
  mutate(mo_day = format(sampleDate, "%m-%d"),
         site_year = paste0(SiteID, year),
         is_19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0)) %>% 
  filter(month(sampleDate) %in% 6:9) %>%
  complete(SiteID, sampleDate) %>% 
  filter(!(grepl("kijilo", SiteID) & sampleDate == as.Date("2019-09-19"))) %>%
  filter(!is.na(Region)) %>% 
  ggplot() +
  geom_line(data = . %>% filter(is_19 == 0), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year), color = "grey") +
  geom_line(data = . %>% filter(is_19 == 1), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year, color = "red")) +
  geom_abline(aes(intercept = 18, slope = 0), linetype = 2) +
  facet_wrap(~Name) +
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank(), text = element_text(size = 16)) +
  labs(y = "Mean Daily Temperature (°C)")

ggdraw(add_sub(b1, "113 sites with one to 18 years of data from 2001-2020 \n2019 data in red",
               x = 0, hjust = 0))
       

ggsave("output/SWIM_figures/daily temps by huc8.jpeg", 
       width = 9, height = 4.5, units = "in", bg = "white")
```


# Thermal regimes

Map of thermal regime classes within Bristol Bay and by year. Note that thermal regimes are ordered from coldest to warmest 6 < 1 < 2 < 3 < 4 < 5. Class 1 has latest timing of warm temps (e.g high snow inputs), and class 6 has very low variance (e.g. groundwater inputs).

* 2012 and 2013 were high snow years, but 2013 generally had quite warm summer temperatures so it must have been a very warm summer.
* 2015 and 2016 were low snow years
* 2019 was hot

```{r map of TR classes for specific years, fig.width = 8}
# TR saved for Dean in AKSSF repo
tr_grps <- read_csv(file = "data/Thermal_regimes_May22.csv")

# left_join(temp %>% mutate(Year = year(sampleDate)), tr_grps) #%>% 

tr_sf <- left_join(md_sf, tr_grps, by = c("Site" = "SiteID", "Region" = "Region"))  

# tr_sf %>% 
#   st_drop_geometry() %>% 
#   count(Year, grp_6) %>% 
#   pivot_wider(names_from = grp_6, values_from = n, names_sort = TRUE)

tr_sf %>% filter(Region == "Bristol Bay", Year %in% c(2012, 2013, 2015, 2016, 2019)) %>% 
  ggplot() +
  geom_sf(data = huc8_wgs84 %>% filter(Name %in% bb_huc8_names)) +
  geom_sf(aes(color = as.factor(grp_6))) +
  facet_wrap(~Year, ncol = 2) +
  theme_bw() +
  labs(color = "Thermal Regime", Title = "Thermal Regimes by Year Across Sites in Bristol Bay")
```



To do plotting of a pca biplot in ggplot, followed these instructions: https://clauswilke.com/blog/2020/09/07/pca-tidyverse-style/


Look at the data in PC coordinates.
Look at the rotation matrix.
Look at the variance explained by each PC.

```{r pca biplot of thermal regimes}

mets11 <- readRDS(file = "data/temp_mets11_wTR.rds")


mets_pca <- prcomp(mets11 %>% filter(Region == "Bristol Bay") %>% select(MWMT:`Tim. MWMT`) , scale = TRUE) 
biplot(mets_pca)
#ggplot version of pca biplot (so can add black circles around sites in BB)


# Get eigenvalues (variance explained by each PC)
mets_pca %>%
  tidy(matrix = "eigenvalues") %>%
  ggplot(aes(PC, percent)) +
  geom_col(fill = "#56B4E9", alpha = 0.8) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  ) +
  theme_minimal_hgrid(12)


# first plot is just points
p1 <- mets_pca %>%
  augment(mets11 %>% filter(Region == "Bristol Bay")) %>% # add original dataset back in
  ggplot()  +
  geom_point(aes(.fittedPC1, .fittedPC2, color = as.factor(grp_6)), size = 1.5) +
  geom_point(aes(.fittedPC1, .fittedPC2), size = 2, shape = 1, data = . %>% filter(Year == 2019)) + 
  # stat_ellipse(aes(.fittedPC1, .fittedPC2, color = as.factor(grp_6))) +
  theme_half_open(12) + 
  background_grid() +
  labs(x = "PC1 (53.5%)", y = "PC2 (17.0%)", color = "Thermal Regimes") +
  theme(legend.position = "bottom")

# second plot is covariates as arrows
arrow_style <- arrow(
  angle = 20, ends = "first", type = "closed", length = grid::unit(8, "pt")
)

# plot rotation matrix
p2 <- mets_pca %>%
  tidy(matrix = "rotation") %>%
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>%
  ggplot(aes(PC1, PC2)) +
  geom_segment(xend = 0, yend = 0, arrow = arrow_style) +
  geom_text_repel(
    aes(label = column), 
    color = "#904C2F"
  ) +
  # xlim(-0.2, .6) + ylim(-.7, 0.3) +
  coord_fixed() + # fix aspect ratio to 1:1
  theme_half_open(12) + 
  background_grid() 

plot_grid(p1, p2)
```

# Thermal sensitivities in Bristol Bay

```{r thermal sensitivity summary}
names(mod_dat)

bb_ts <- mod_dat %>% 
  select(Region:TempSens_Air.DayLen) %>% 
  filter(Region == "Bristol Bay") %>% 
  group_by(Region, Site) %>% 
  summarize(mn_ts = mean(TempSens_Air.DayLen),
            sd_ts = sd(TempSens_Air.DayLen),
            yr_ct = n())


ggplot() +
  geom_point(aes(x = mn_ts, y = sd_ts), data = bb_ts)
```

```{r temp Sens map}
ggplot() +
  geom_sf(data = huc8_wgs84 %>% right_join(bb_huc8_names), aes(fill = Name)) +
  geom_sf(data = md_sf %>% right_join(bb_ts) %>% filter(yr_ct > 2), aes(size = sd_ts), shape = 1)
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(title = "Bristol Bay Long-Term Sites", subtitle = "10+ years of data") +
  guides(fill = "none")

# ggsave("output/SWIM_figures/map of tempSens.jpeg", width = 9, height = 4, units = "in")

```


```{r}
left_join(bb_ts, md_sf %>% select(Site, Name)) %>% 
  filter(yr_ct > 2) %>% 
  ggplot() +
  geom_histogram(aes(x = mn_ts)) 
  

```


# Trends in metrics for long-term sites


Look for sites with complete summers of data and 10 or more years.

```{r long-term sites}
bb_ltsites <- bb_temp %>% 
  filter(month(sampleDate) %in% 6:8) %>% 
  group_by(SiteID, year) %>% 
  mutate(yr_ct = n()) %>% 
  filter(yr_ct > 73) %>% 
  distinct(SiteID, year) %>% 
  ungroup() %>% 
  group_by(SiteID) %>% 
  mutate(yr_ct = n()) %>%
  filter(yr_ct > 9)

bb_ltdat <- left_join(bb_ltsites, bb_temp)

bb_ltsites %>% distinct(SiteID)
```


```{r map of lt sites}

ggplot() +
  geom_sf(data = huc8_wgs84 %>% right_join(bb_huc8_names), aes(fill = name_lab)) +
  geom_sf(data = md_sf %>% right_join(bb_ltsites %>% distinct(SiteID), by = c("Site" = "SiteID"))) +
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(title = "Bristol Bay Long-Term Sites", subtitle = "10+ years of data") +
  guides(fill = "none")

ggsave("output/SWIM_figures/map of BB lt sites.jpeg", width = 9, height = 4, units = "in")

```


```{r summer metrics for lt sites}

bb_ltmets <- bb_ltdat %>% 
  filter(month(sampleDate) %in% 6:8) %>%
  mutate(jd = format(sampleDate, "%j") %>% as.numeric()) %>% 
  group_by(SiteID, year) %>% 
  mutate(mn_7day = rollapply(meanDT, 7, mean, fill = NA),
         max_7day = max(mn_7day, na.rm = TRUE),
         mwat_jd = case_when(mn_7day == max_7day ~ jd,
                             TRUE ~ 0)) %>% 
  # select(SiteID, year, jd, meanDT, mn_7day, max_7day, mwat_jd)
  summarize(MWAT = max(mn_7day, na.rm = TRUE),
            Summer = mean(meanDT),
            MWAT_jd = max(mwat_jd))

summary(bb_ltmets)
```



```{r mann kendall trend test}

bb_mktests <- bb_ltmets %>%
  pivot_longer(cols = MWAT:MWAT_jd) %>% 
  group_by(SiteID, name) %>%  
  nest() %>%
  mutate(
    mk.test = map(data, ~ mk.test(.x$value)),
    sen.test = map(data, ~ sens.slope(.x$value)),
    mk.tidy = map(mk.test, tidy),
    sen.tidy = map(sen.test, tidy),
    sen.est = map(sen.test, "estimates"))
    
bb_mkres <- bind_rows(bb_mktests%>% 
            unnest(mk.tidy) %>% 
              select(SiteID, name, statistic, p.value, method), 
          bb_mktests %>% 
            unnest(sen.tidy) %>% 
              select(SiteID, name, statistic, p.value, method)) %>% 
  left_join(bb_mktests %>% unnest(sen.est) %>% select(SiteID, name, sen.est)) %>% 
  arrange(p.value)

bb_mkres %>% 
  filter(p.value < 0.05) %>% 
  arrange(name, sen.est)

```



```{r trends for lt sites}
bb_ltmets %>% 
  pivot_longer(cols = MWAT:MWAT_jd) %>% 
  ggplot(aes(x = year, y = value)) +
  geom_line(aes(group = SiteID), color = "gray") +
  geom_smooth(method = "lm") +
  facet_wrap(~name, scales = "free") +
  theme_bw() +
  theme()
```

```{r significant trends}

bb_ltmets %>%
  left_join(md %>% select(Site, Waterbody_name), by = c("SiteID" = "Site")) %>%
  mutate(wb_name = case_when(is.na(Waterbody_name) ~ "Aleknagik Bear Creak",
                             Waterbody_name == "Lynx" ~ "Nerka Lynx Creek",
                             TRUE ~ str_to_title(Waterbody_name))) %>% 
  pivot_longer(cols = MWAT:MWAT_jd) %>% 
  right_join(bb_mkres %>% filter(p.value < 0.05) %>% select(SiteID, name)) %>% 
  ggplot(aes(x = year, y = value, color = wb_name)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~name, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = 16), legend.position = "bottom") +
  labs(color = "", x = "Year", y = "Metric Value") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))

ggsave("output/SWIM_figures/temp trends.jpeg", width = 7, height = 4, units = "in")
```

Recreate above plot but show all temporal patterns with the significant trends highlighted.

```{r all trends in gray}

bb_ltmets %>%
  left_join(md %>% select(Site, Waterbody_name), by = c("SiteID" = "Site")) %>%
  mutate(wb_name = case_when(is.na(Waterbody_name) ~ "Aleknagik Bear Creak",
                             Waterbody_name == "Lynx" ~ "Nerka Lynx Creek",
                             TRUE ~ str_to_title(Waterbody_name))) %>% 
  pivot_longer(cols = MWAT:MWAT_jd) %>% 
  left_join(bb_mkres %>% filter(p.value < 0.05) %>% select(SiteID, name) %>% 
               mutate(sig_trend = 1)) %>%
  mutate(name_lab = factor(name, levels = c("Summer", "MWAT", "MWAT_jd"),
                           labels = c("Mean Summer Temperature (Celsius)", "Max Weekly Temperature (Celsius)", "Timing of Max Weekly Temperature (Julian Day)"))) %>% 
  ggplot() +
  # geom_point(aes(x = year, y = value, group = SiteID), color = "gray") +
  geom_line(aes(x = year, y = value, color = SiteID)) +
  facet_wrap(~name_lab, scales = "free", strip.position="left") +
  theme_bw() +
  theme(text = element_text(size = 16), legend.position = "bottom") + guides(colour = FALSE)+
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        # Customize theme so that is black & white style as requested
        panel.background = element_rect(fill = NA, colour = 'black'),
        panel.grid = element_blank())+
  labs(color = "", x = "Year", y = NULL) +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))

ggsave("output/SWIM_figures/all temp trends.jpeg", width = 9, height = 6, units = "in")
```

```{r all trends with sig trends highlighted}

bb_ltmets %>%
  left_join(md %>% select(Site, Waterbody_name), by = c("SiteID" = "Site")) %>%
  mutate(wb_name = case_when(is.na(Waterbody_name) ~ "Aleknagik Bear Creak",
                             Waterbody_name == "Lynx" ~ "Nerka Lynx Creek",
                             TRUE ~ str_to_title(Waterbody_name))) %>% 
  pivot_longer(cols = MWAT:MWAT_jd) %>% 
  left_join(bb_mkres %>% filter(p.value < 0.05) %>% select(SiteID, name) %>% 
               mutate(sig_trend = 1)) %>%
  mutate(name_lab = factor(name, levels = c("Summer", "MWAT", "MWAT_jd"),
                           labels = c("Mean Summer Temp.", "Max Weekly Temp.", "Timing of Max Weekly Temp."))) %>% 
  ggplot() +
  # geom_point(aes(x = year, y = value, group = SiteID), color = "gray") +
  geom_line(aes(x = year, y = value, group = SiteID), color = "gray") +
  geom_point(aes(x = year, y = value, color = wb_name),
             data = . %>% filter(sig_trend == 1)) +
  geom_line(aes(x = year, y = value, color = wb_name),
             data = . %>% filter(sig_trend == 1)) +
  geom_smooth(aes(x = year, y = value, color = wb_name), 
              data = . %>% filter(sig_trend == 1), method = "lm") +
  facet_wrap(~name_lab, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = 16), legend.position = "bottom") +
  labs(color = "", x = "Year", y = "Metric Value") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))

ggsave("output/SWIM_figures/temp trends.jpeg", width = 9, height = 4, units = "in")
```
#Looking at temperature metrics for sites
```{r}
head(mets)

#filtering for BB sites only - need to merge with site metadata first
BB_mets<-mets%>%
  left_join(md, by=c("SiteID"="Site"))%>%
  filter(Region=="Bristol Bay")
dim(BB_mets)

#making figure similar to Figure 4 in Mauger 2016 of daily occurrence of MWMT 
ggplot(data=BB_mets, aes(x=as.Date(MA7d_DMT_jd, origin="1970-01-01")))+geom_histogram(aes(fill=Year), binwidth=1)+theme_bw()+xlab("Month")+ylab("MWMT Count by Day")

#faceting on year - just out of curiousity, unlikely to use this in reporting since it's too messy
ggplot(data=BB_mets, aes(x=as.Date(MA7d_DMT_jd, origin="1970-01-01")))+geom_histogram(aes(color=Year), binwidth=1)+theme_bw()+xlab("Month")+ylab("MWMT Count by Day")+facet_wrap(~Year)

#making plots of MWMT, mean temp, days >18 for all AKSSF sites
keep_cols<-c("SiteID", "Year", "Region", "HUC8", "MA7d_DMT", "MnDAT", "SUM_18_DMT")

all_mets<-mets%>%
  left_join(md, by=c("SiteID"="Site"))%>%
  select(keep_cols)

head(all_mets)

all_mets_long<-all_mets%>%
  pivot_longer(!c("SiteID", "Year", "Region", "HUC8"), names_to="metric", values_to="values")%>%
  mutate(metricf=factor(metric, levels=c("MA7d_DMT", "MnDAT", "SUM_18_DMT"), 
                         labels=c("Maximum Weekly Mean Temperature (Celsius)", "Mean Temperature (Celsius)", expression("Number of Days > 18°C"))))

ggplot(all_mets_long, aes(x=fct_rev(Region), y=values))+xlab("Region")+geom_boxplot()+facet_wrap(~metricf, scales="free", strip.position="bottom", ncol=1)+theme_bw()+coord_flip()+ylab(NULL)
ggsave("output/SWIM_figures/all region boxplots.jpg", width=5, height=10, units="in", dpi=1000)

#making plots of MWMT, mean temp, days >18 for just BB sites by HUC


huc8_names<-huc8%>%
  select("Name", "HUC8")%>%
  as.data.frame()

BB_huc_metrics<-all_mets_long%>%
  mutate(HUC8=as.character(HUC8))%>%
  left_join(huc8_names)%>%
  filter(Region=="Bristol Bay")

ggplot(BB_huc_metrics, aes(x=Name, y=values))+xlab("Watershed")+geom_boxplot()+facet_wrap(~metricf, scales="free", strip.position="bottom", ncol=1)+theme_bw()+coord_flip()+ylab(NULL)

ggsave("output/SWIM_figures/Bristol Bay HUC8 boxplots.jpg", width=5, height=10, units="in", dpi=1000)


```

#making PCA biplot of Bristol Bay thermal regimes by HUC8
```{r pca biplot of BB Huc8 thermal regimes}

pca_keepcols <- c("SiteID", "Year", "Name", "MA7d_DMT", "RANGE_MAX", "SIGMA_DMT", "SUM_13_DMT", "SUM_18_DMT", "dur13mx", "dur18mx",
             "MxDMT_jd", "MA7d_DMT_jd", "NDNT", "MnDAT")

huc_mets<-mets%>%
  left_join(md, by=c("SiteID"="Site"))%>%
   mutate(HUC8=as.character(HUC8))%>%
  left_join(huc8_names)%>%
  filter(Region=="Bristol Bay")%>%
  select(all_of(pca_keepcols))

hucmets_lg <- huc_mets %>% 
  pivot_longer(cols = MA7d_DMT:MnDAT, names_to = "metric", values_to = "value") %>% 
  mutate(group = case_when(grepl("SIGMA|MxDIFF|SD|CV|RANGE", metric) ~ "Variability",
                           grepl("dur", metric) ~ "Duration",
                           grepl("SUM", metric) ~ "Frequency",
                           grepl("jd", metric) ~ "Timing",
                           grepl("TempSens", metric) ~ "Sensitivity",
                           TRUE ~ "Magnitude"))

mets11_scwd <- hucmets_lg %>%
  mutate(metric_lab = case_when(metric == "MA7d_DMT" ~ "MWMT",
                                metric == "RANGE_MAX" ~ "Range",
                                metric == "SIGMA_DMT" ~ "Var.",
                                metric == "SUM_13_DMT" ~ "Sum 13",
                                metric == "SUM_18_DMT" ~ "Sum 18",
                                metric == "dur13mx" ~ "Dur. 13",
                                metric == "dur18mx" ~ "Dur. 18",
                                metric == "MxDMT_jd" ~ "Tim. Max. Daily",
                                metric == "MA7d_DMT_jd" ~ "Tim. MWMT",
                                metric == "NDNT" ~ "Min. Temp.",
                                metric == "MnDAT" ~ "Mean Temp.")) %>% 
  group_by(metric_lab) %>% 
  mutate(value_sc = scale(value)) %>% 
  select(SiteID, Year, Name, metric_lab, value_sc) %>% 
  pivot_wider(names_from = metric_lab, values_from = value_sc)


mets11_dist <- dist(as.data.frame(mets11_scwd) %>% select('MWMT':'Mean Temp.'))

mets11_clust <- hclust(mets11_dist, method = "ward.D2") 

mets11_tr <- bind_cols(mets11_scwd, 
                       cutree(mets11_clust, k = 3:10) %>% 
                         as_tibble() %>% 
                         rename_with(.fn = ~ paste0("grp_", .x)))


bind_cols(mets11_scwd, group = cutree(mets11_clust, k = 6)) %>% 
  ggplot(aes(x = as.factor(group), fill = Name)) +
  geom_bar(position = "fill")

bind_cols(mets11_scwd, group = cutree(mets11_clust, k = 6)) %>% 
  count(group, Name) %>%
  group_by(Name) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(x = as.factor(group), y = percent, fill = HUC8)) +
  geom_col(position = position_dodge()) 


pc1 <- prcomp(mets11_scwd %>% select(-(SiteID:Name))) 

pc2 <- princomp(mets11_scwd %>% select(-(SiteID:Name)), cor = TRUE) 

biplot(pc1)
biplot(pc2)


autoplot(pc2, data = mets11_scwd, color="Name", labels = TRUE, loadings.label = TRUE,loadings.label.colour = "black", loadings.label.repel = TRUE, loadings.colour = "black") +
  theme_bw()

autoplot(pc2, data = mets11_scwd, frame.type='norm', fill=NULL, frame.colour = "Name") +
  theme_bw()

```

