---
title: "3_fish_temps"
output: 
  html_document:
    code_folding: hide
date: "2023-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(zoo)
library(lubridate)
```

This file is for exploring ST at the sites where UW has been monitoring coho salmon sizes over time. Figures are shown in this markdown and saved to the temperature/output folder for use in presentations. Currently planning an update at the 2023 SWIM meeting.

```{r read in temps}
temp <- read_rds("fish/data/temp_data_2023-03-16.rds")

temp %>% 
  count(StationName, YearSampled) %>% 
  pivot_wider(names_from = YearSampled, values_from = n, names_sort = TRUE) %>% 
  kable()
```

```{r plot of ST at different sites and years}

temp %>% 
  mutate(day = format(sampleDate, "%m-%d")) %>% 
  ggplot(aes(x = as.Date(day, format = "%m-%d"), y = meanDT)) + 
  geom_line(aes(color = as.factor(YearSampled))) +
  facet_wrap(~StationName) +
  theme_bw() +
  theme(axis.title.x = element_blank(), text = element_text(size = 16)) +
  labs(color = "Year", y = "Mean Daily Temperature (°C)",
       title = "Stream Temperatures at UW Monitoring Sites")


ggsave("temperature/output/UW_fish_sites/Daily temps for 8 sites.jpeg")
```

Calculating annual stream temperature metrics for Erik:

* CDD for all days in June-August
* CDD for June 21st (julian day 172) through end of August so sum is across a consistent set of days with data for all sites
* mean summer temperature - common metric used by UW to compare stream thermal regimes
* MWAT - maximum weekly average temperature. Maximum of 7-day rolling average of mean daily temperatures 

```{r calculation of annual temperature metrics}

# one site-year that started on jd 172.
temp %>% 
  mutate(jd = format(sampleDate, "%j") %>% as.numeric()) %>% 
  group_by(StationName, YearSampled) %>% 
  summarize(minjd = min(jd))


temp_mets <- temp %>% 
  filter(month(sampleDate) %in% 6:8) %>%
  group_by(StationName, YearSampled) %>% 
  mutate(jd = format(sampleDate, "%j") %>% as.numeric(),
         meanDT_172 = case_when(jd < 172 ~ 0,
                                TRUE ~ meanDT),
         cumtemp = cumsum(meanDT),
         cumtemp_172 = cumsum(meanDT_172),
         mn_7day = rollapply(meanDT, 7, mean, fill = NA)) %>% 
  summarize(CDD = max(cumtemp),
            CDD_172 = max(cumtemp_172),
            MWAT = max(mn_7day, na.rm = TRUE),
            mn_summer = mean(meanDT),
            count = n())
  
summary(temp_mets)

saveRDS(temp_mets, paste0("fish/data/temp_metrics_", Sys.Date(), ".rds"))
```


Figures show the mean and range for each metric. Once we have models or data to fill in all years at all sites, we could do a bootstrapped confidence interval on the mean (mean_cl_boot to stat_summary). 

```{r CDD figure}

temp_mets %>% 
  group_by(StationName) %>% 
  mutate(yr_ct = n(),
         site_lab = paste0(gsub(" Creek", "", StationName), " (", yr_ct, ")")) %>% 
  ggplot(aes(y = fct_reorder(site_lab, CDD_172), x = CDD_172)) +
  stat_summary(fun = mean, fun.min = min, fun.max = max, color = "black") +
  # geom_point(color = "gray") +
  theme_bw() +
  theme(axis.title.y = element_blank(), text = element_text(size = 16)) +
  labs(color = "Year", x = "CDD (°C)",
       title = "UW Monitoring Sites",
       subtitle = "Cumulative Degree Days from June 20 - August 31")

ggsave("temperature/output/UW_fish_sites/CDD for 8 sites.jpeg")

temp_mets %>% 
  ggplot(aes(x = YearSampled, y = CDD_172, color = StationName)) +
  geom_line() +
  geom_point()

```


```{r MWAT figure}

temp_mets %>% 
  group_by(StationName) %>% 
  mutate(yr_ct = n(),
         site_lab = paste0(gsub(" Creek", "", StationName), " (", yr_ct, ")")) %>% 
  ggplot(aes(y = fct_reorder(site_lab, MWAT), x = MWAT)) +
  stat_summary(fun = mean, fun.min = min, fun.max = max, color = "black") +
  # geom_point(color = "gray") +
  theme_bw() +
  theme(axis.title.y = element_blank(), text = element_text(size = 16)) +
  labs(color = "Year", x = "MWAT (°C)",
       title = "UW Monitoring Sites",
       subtitle = "Maximum Weekly Average Stream Temperature")

ggsave("temperature/output/UW_fish_sites/MWAT for 8 sites.jpeg")


```
