---
title: "CleanPaperAnalysis"
author: "Erin Larson"
date: "2025-03-07"
output: html_document
---
Got Egegik data added in - check in AKTEMP for other sites that might be missing data???
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(sf)
library(lubridate)
library(broom)
library(cowplot)
library(ggrepel)
library(zoo)
library(trend)
library(Kendall)
library(ggfortify)
library(RColorBrewer)
library(ggspatial)
library(GSODR)
library(rstudioapi)
```

# Read in data and add attributes

All data were combined in the AKSSF repository. Read from that project drive and filter on data from Bristol Bay.

NOTE: Data from Zenodo are incomplete and are missing the UW data from 2017 on.

```{r DFA input data}

temp <- read_csv("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\summer_data_wair_dayl2022-03-17.csv")
md <- read_csv("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\md_2022-02-08.csv")

md %>% count(Waterbody_type)
md %>% filter(Region == "Bristol Bay") %>% distinct(SourceName, Contact_person)
```

Adding watershed attributes created for the AKSSF project to these sites.

```{r read in TS model data frame}
mod_dat <- readRDS("C:\\Users\\elarson15\\Documents\\Projects\\229904 AKSSF Thermal Sensitivity\\AKSSF\\data_preparation\\final_data\\model_data2022-05-09.rds")

mod_dat


##accessing AKTEMP to add in 2019 and 2020 data for Egegik Bay - not included in the original dataset for some reason...
Egegik<-read.csv("data/FWS/AKTEMP-USFWS_EgegikR_daily.csv")
head(Egegik)
Egegik$series_id<-NULL
Egegik$flag<-NULL
Egegik$n_values<-NULL
Egegik$SiteID<-rep("fws_580223156504200")
Egegik$sampleDate<-as.Date(Egegik$date, format="%m/%d/%Y")
Egegik$date<-NULL
Egegik$year<-year(Egegik$sampleDate)
colnames(Egegik)<-c("minDT", "meanDT", "maxDT", "SiteID", "sampleDate", "year")
study_period<-filter(Egegik, sampleDate<"2020-10-01" & sampleDate>"2018-10-01")

all_temp<-bind_rows(temp, study_period)

#note that Tim had me populate all airtemps for this data frame, can remove dates without meandt
bb_temp <- left_join(all_temp, md %>% 
                       select(Site, SourceName, Waterbody_name, Waterbody_type, Region, Name, Latitude, Longitude, Waterbody_name, HUC8), by = c("SiteID" = "Site")) %>% 
    filter(Region == "Bristol Bay", !is.na(meanDT)) 

#9 sites in temperature data frame that we didn't calculate geospatial data for.
# I think these were dropped bc too little data or incomplete summer data.
anti_join(bb_temp %>% distinct(SiteID), mod_dat %>% distinct(Site), by = c("SiteID" = "Site"))

names(mod_dat)

bb_temp <- left_join(bb_temp, mod_dat %>% 
            select(Site, str_ord:wtd_area_sqKM) %>% 
            distinct(), 
          by = c("SiteID" = "Site")) %>% 
  filter(!is.na(str_ord))



```



```{r BB data summary for Will and Erik}

#creating site/date/mean daily temp frame for Erik and Will
mean_dailies<-bb_temp%>%
  select(SiteID, Latitude, Longitude, HUC8, Waterbody_name, sampleDate, meanDT, str_ord)

write.csv(mean_dailies, file="output/BBDailyTemps.csv")


#making table for Will for map
bb_years<-bb_temp %>% 
  filter(Region == "Bristol Bay") %>% 
  distinct(SiteID, year = year(sampleDate)) %>% 
  count(SiteID) %>% 
  arrange(n)%>%
  rename(NumberYears=n)%>%
  left_join(y=distinct(bb_temp, SiteID, Latitude, Longitude))

head(bb_years)

write.csv(bb_years, "output/BBThermalRegimeSiteInfoForMapping.csv")  
            
            

```


# Site summaries

## Map

Map of sites for each HUC8 in Bristol Bay. (Looks like the metadata I created already has HUC8 on there.)

```{r BB sites by huc8}

bbtemp_sf <- st_as_sf(bb_temp, coords = c("Longitude", "Latitude"), crs = "WGS84")

huc8 <- st_read(dsn = "S:\\GIS_Layers\\GIS_data\\inlandWaters\\Hydrography\\NHD\\WBD\\WBD_19_GDB.gdb", layer = "WBDHU8")

huc8_wgs84 <- st_transform(huc8, crs = "WGS84")



bb_huc8_names <- bbtemp_sf %>%
    st_drop_geometry() %>% 
   filter(Region == "Bristol Bay") %>% 
  distinct(SiteID, year = year(sampleDate), Name) %>% 
  count(SiteID) %>% 
  arrange(n)%>%
  rename(NumberYears=n)%>%
  left_join(y=distinct(bb_temp, SiteID, Latitude, Longitude, Name))%>%
  count(Name) %>% 
  mutate(name_lab = paste0(gsub("River", "R.", Name), " (", n, ")"))

locations.sf<-st_as_sf(bbtemp_sf )%>%
ggplot()+
  geom_sf(data = huc8_wgs84 %>% right_join(bb_huc8_names), aes(fill = name_lab)) +
  geom_sf(data = bbtemp_sf ) +
  theme_bw() +
  theme(text = element_text(size = 18)) +
  labs(fill = "HUC8 Name", title = "Bristol Bay Sampling Sites")

write.csv(md %>% filter(Region == "Bristol Bay"), "output/TempMonitoringLocations.csv")
ggsave("output/SWIM_figures/map of BB sites.jpeg", width = 9, height = 4, units = "in")


```


## Sites by watershed attributes and years


What sites have been monitored or are being monitored?

- number of sites in each HUC8 (possibly filter to exclude those with < 3 years of data): sites by huc8 code chunk
- number of years of data for each site: site count by years of data code chunk

Do sites represent the different stream sizes/types in Bristol Bay?

- sites by stream order or elevation.
- compare this histogram to Dustin's summary of stream order or elevation for all HUC12 outlets: sites by stream order code chunk

Do sites represent different thermal regimes?

- look at BB sites in pca biplot colored by different TR: pca biplot of thermal regimes code chunk

```{r site count by years of data}
bb_temp %>% 
  filter(month(sampleDate) %in% 6:8) %>% 
  group_by(SiteID, year) %>% 
  mutate(yr_ct = n()) %>% 
  filter(yr_ct > 73) %>% 
  distinct(SiteID, year) %>% 
  ungroup() %>% 
  count(year) %>%
  # arrange(desc(n))
  ggplot(aes(x = year, y = n)) +
  geom_col() +
  theme_bw() +
  theme() +
  labs(x = "Year", y = "Count of Sites")

ggsave("output/SWIM_figures/site count by year.jpeg", width = 4.5, height = 4.5, units = "in")
```


```{r sites by stream order}
bb_temp %>% 
  filter(str_ord > 5) %>% 
  distinct(SiteID, str_ord) %>% 
  filter(!is.na(str_ord)) %>% 
  count(str_ord)

#read in Dustin's data frame with stream order for all huc12 outlets
bb_huc12 <- read_csv("data/spatial/AKSSF_AwcHuc12s_strwtd_cv.csv") %>% 
  filter(region == "Bristol_Bay")
names(bb_huc12)  

bb_huc12 %>%
  count(str_ord)

bind_rows(bb_huc12 %>% select(SiteID = cat_ID_con, str_ord) %>% mutate(type = "HUC12"),
          bb_temp %>% select(SiteID, str_ord) %>% mutate(type = "Monitoring Sites")) %>% 
  ggplot() +
  geom_histogram(aes(x = (str_ord), 
                     y=stat(density*width), fill = type), 
                 position = position_dodge()) 

bind_rows(bb_huc12 %>% select(SiteID = cat_ID_con, str_ord) %>% mutate(type = "HUC12 Outlets"),
          bb_temp %>% select(SiteID, str_ord) %>% mutate(type = "Monitoring Sites")) %>%
  count(type, order = factor(str_ord)) %>% 
  group_by(type) %>% 
  mutate(n = prop.table(n)) %>% 
  ggplot(aes(x = order, y = n, fill = type)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Stream Order", y = "Percent of Sites", fill = "Site Type")

ggsave("output/SWIM_figures/site percent by stream order.jpeg", width = 4.5, height = 4.5, units = "in")

```




# Time series figures

Plots of daily temps by HUC8 and with 20C threshold for thermal stress to adults and juveniles.



```{r time series by huc8}

#Making timeseries by strahler order

b1 <- bb_temp %>% 
  mutate(mo_day = format(sampleDate, "%m-%d"),
         site_year = paste0(SiteID, year),
         is_19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0),
         large_drainage=case_when(
           Name=="Mulchatna River"~"Nushagak River",
           Name=="Egegik Bay"~"Egegik Bay",
           Name=="Lake Clark"~"Kvichak River",
           Name=="Lake Iliamna"~"Kvichak River",
           Name=="Lower Nushagak River"~"Nushagak River",
           Name=="Naknek"~"Naknek River",
           Name=="Nushagak Bay"~"Nushagak Bay",
           Name=="Togiak"~"Togiak Bay",
           Name=="Upper Nushagak River"~"Nushagak River",
           Name=="Wood River"~"Wood River",
         )) %>% 
  filter(month(sampleDate) %in% 6:9) %>%
    filter(year(sampleDate) %in% 2015:2019) %>%
  complete(SiteID, sampleDate) %>% 
  filter(!(grepl("kijilo", SiteID) & sampleDate == as.Date("2019-09-19"))) %>%
  filter(!is.na(Region)) %>% 
  ggplot() +
  geom_line(data = . %>% filter(is_19 == 0), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year), color = "grey") +
  geom_line(data = . %>% filter(is_19 == 1), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year, color = "red")) +
  geom_abline(aes(intercept = 20, slope = 0), linetype = 2) +
  facet_wrap(~str_ord) +
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank(), text = element_text(size = 16)) +
  labs(y = "Mean Daily Temperature (°C)")

b1

ggsave("output/SWIM_figures/daily temps by stream order_2015_2019.jpeg", 
       width = 9, height = 4.5, units = "in", bg = "white")


#Making new timeseries with just 2015-2019 data

b2 <- bb_temp %>% 
  mutate(mo_day = format(sampleDate, "%m-%d"),
         site_year = paste0(SiteID, year),
         is_19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0),
         large_drainage=case_when(
           Name=="Mulchatna River"~"Nushagak River",
           Name=="Egegik Bay"~"Egegik Bay",
           Name=="Lake Clark"~"Kvichak River",
           Name=="Lake Iliamna"~"Kvichak River",
           Name=="Lower Nushagak River"~"Nushagak River",
           Name=="Naknek"~"Naknek River",
           Name=="Nushagak Bay"~"Nushagak Bay",
           Name=="Togiak"~"Togiak Bay",
           Name=="Upper Nushagak River"~"Nushagak River",
           Name=="Wood River"~"Wood River",
         )) %>% 
  filter(month(sampleDate) %in% 6:9) %>%
    filter(year(sampleDate) %in% 2015:2019) %>%
  complete(SiteID, sampleDate) %>% 
  filter(!(grepl("kijilo", SiteID) & sampleDate == as.Date("2019-09-19"))) %>%
  filter(!is.na(Region)) %>% 
  ggplot() +
  geom_line(data = . %>% filter(is_19 == 0), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year), color = "grey") +
  geom_line(data = . %>% filter(is_19 == 1), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year, color = "red")) +
  geom_abline(aes(intercept = 20, slope = 0), linetype = 2) +
  facet_wrap(~factor(large_drainage, level=c("Togiak Bay", "Nushagak Bay", "Wood River", "Nushagak River", "Kvichak River", "Naknek River", "Egegik Bay"))) +
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank(), text = element_text(size = 16)) +
  labs(y = "Mean Daily Temperature (°C)")

b2

ggsave("output/SWIM_figures/daily temps by largedrainage_2015_2019.jpeg", 
       width = 9, height = 4.5, units = "in", bg = "white")

```
#Making histograms of MWMT by DOY, based on drainage and year
```{r}
head(mets)

bb_strahler<-bb_temp%>%
  select(SiteID, str_ord)%>%
  distinct()

all_md<-md%>%
    left_join(bb_strahler, by=c("Site"="SiteID"))

#filtering for BB sites only - need to merge with site metadata first
BB_mets<-mets%>%
  left_join(all_md, by=c("SiteID"="Site"))%>%
  filter(Region=="Bristol Bay")

dim(BB_mets)

BB_mets_summary<-BB_mets%>%filter(Year %in% 2015:2019)%>%
         mutate(  #this command tells R to create a new column based on the HUC8 names to create the new large drainage names 
         large_drainage=case_when(
           Name=="Mulchatna River"~"Nushagak River",
           Name=="Egegik Bay"~"Egegik Bay",
           Name=="Lake Clark"~"Kvichak River",
           Name=="Lake Iliamna"~"Kvichak River",
           Name=="Lower Nushagak River"~"Nushagak River",
           Name=="Naknek"~"Naknek River",
           Name=="Nushagak Bay"~"Nushagak Bay",
           Name=="Togiak"~"Togiak Bay",
           Name=="Upper Nushagak River"~"Nushagak River",
           Name=="Wood River"~"Wood River",
         ))

#making figure similar to Figure 4 in Mauger 2016 of daily occurrence of MWMT w/ fill by drainage
ggplot(data=BB_mets_summary, aes(x=as.Date(MA7d_DMT_jd, origin="1970-01-01"), fill=large_drainage))+geom_histogram( binwidth=1)+theme_bw()+xlab("Month")+ylab("MWMT Count by Day")+ scale_fill_brewer(type="div",palette="Dark2", guide=guide_legend(reverse=TRUE), name="Drainage", breaks=c("Egegik Bay", "Naknek River", "Kvichak River", "Nushagak River","Wood River", "Nushagak Bay","Togiak Bay"))

ggsave("output/SWIM_figures/MWMTByDay_Drainage.jpg", width=5, height=3, units="in", dpi=1000)


#making figure similar to Figure 4 in Mauger 2016 of daily occurrence of MWMT w/ fill by stream order
ggplot(data=BB_mets_summary%>%filter(str_ord %in% 1:9), aes(x=as.Date(MA7d_DMT_jd, origin="1970-01-01"), fill=as.factor(str_ord)))+geom_histogram( binwidth=1)+theme_bw()+xlab("Month")+ylab("MWMT Count by Day")+scale_fill_brewer(type="seq", palette="Blues", name="Strahler Order")
ggsave("output/SWIM_figures/MWMTByDay_StrahlerOrder.jpg", width=5, height=3, units="in", dpi=1000)


#making figure similar to Figure 4 in Mauger 2016 of daily occurrence of MWMT w/ fill by year
ggplot(data=BB_mets_summary, aes(x=as.Date(MA7d_DMT_jd, origin="1970-01-01"), fill=as.factor(Year)))+geom_histogram( binwidth=1, color="black")+theme_bw()+xlab("Month")+ylab("MWMT Count by Day")+ scale_fill_brewer(type="seq",palette="YlGnBu", name="Year")

ggsave("output/SWIM_figures/MWMTByDay_Year.jpg", width=5, height=3, units="in", dpi=1000)
```


#Making temperature boxplots by HUC8 and drainage
```{r}

#making plots of MWMT, mean temp, days >18 for all AKSSF sites
keep_cols<-c("SiteID", "Year", "Region", "HUC8", "MA7d_DMT", "MnDAT", "SUM_18_DMT", "str_ord")

all_mets<-mets%>%
  left_join(all_md, by=c("SiteID"="Site"))%>%
  select(keep_cols)

head(all_mets)

all_mets_long<-all_mets%>%
  pivot_longer(!c("SiteID", "Year", "Region", "HUC8", "str_ord"), names_to="metric", values_to="values")%>%
  mutate(metricf=factor(metric, levels=c("MA7d_DMT", "MnDAT", "SUM_18_DMT"), 
                         labels=c("Maximum Weekly Mean Temperature (Celsius)", "Mean Temperature (Celsius)", expression("Number of Days > 18°C"))))

ggplot(all_mets_long, aes(x=fct_rev(Region), y=values))+xlab("Region")+geom_boxplot()+facet_wrap(~metricf, scales="free", strip.position="bottom", ncol=1)+theme_bw()+coord_flip()+ylab(NULL)
ggsave("output/SWIM_figures/all region boxplots.jpg", width=5, height=10, units="in", dpi=1000)

#making plots of MWMT, mean temp, days >18 for just BB sites by HUC


huc8_names<-huc8%>%
  select("Name", "HUC8")%>%
  as.data.frame()

BB_huc_metrics<-all_mets_long%>%
  mutate(HUC8=as.character(HUC8))%>%
  left_join(huc8_names)%>%
  filter(Region=="Bristol Bay")

ggplot(BB_huc_metrics, aes(x=Name, y=values))+xlab("Watershed")+geom_boxplot()+facet_wrap(~metricf, scales="free", strip.position="bottom", nrow=1)+theme_bw()+coord_flip()+ylab(NULL)

ggsave("output/SWIM_figures/Bristol Bay HUC8 boxplots.jpg", width=15, height=5, units="in", dpi=1000)


BB_largedrainage_metrics <- BB_huc_metrics %>% 
  mutate(  #this command tells R to create a new column based on the HUC8 names to create the new large drainage names 
         large_drainage=case_when(
           Name=="Mulchatna River"~"Nushagak River",
           Name=="Egegik Bay"~"Egegik Bay",
           Name=="Lake Clark"~"Kvichak River",
           Name=="Lake Iliamna"~"Kvichak River",
           Name=="Lower Nushagak River"~"Nushagak River",
           Name=="Naknek"~"Naknek River",
           Name=="Nushagak Bay"~"Nushagak Bay",
           Name=="Togiak"~"Togiak Bay",
           Name=="Upper Nushagak River"~"Nushagak River",
           Name=="Wood River"~"Wood River",
         ))%>%
  filter(Year %in% 2015:2019)
ggplot(BB_largedrainage_metrics, aes(x=factor(large_drainage, level=c("Egegik Bay", "Naknek River", "Kvichak River", "Nushagak River","Wood River", "Nushagak Bay","Togiak Bay")), #this gets watersheds to display in the correct southeast to northwest order
                                     y=values, color=large_drainage))+ scale_color_brewer(type="div",palette="Dark2")+ #this is the color palette I used
  xlab("Watershed")+geom_boxplot(lwd=0.75)+facet_wrap(~metricf, scales="free", strip.position="bottom", nrow=1)+theme_bw()+coord_flip()+ylab(NULL)+guides(color="none")+theme(text=element_text(size=16))

ggsave("output/SWIM_figures/Bristol Bay Large Drainage boxplots.jpg", width=16, height=5, units="in", dpi=1000)

BB_largedrainage_metrics <- BB_huc_metrics %>%
  filter(Year %in% 2015:2019)

ggplot(BB_largedrainage_metrics, aes(x=str_ord, y=values, color=as.factor(str_ord)))+ scale_color_brewer(type="seq", palette="Blues", name="Strahler Order")+
  xlab("Strahler Order")+geom_boxplot(lwd=0.75)+facet_wrap(~metricf, scales="free", strip.position="bottom", nrow=1)+theme_bw()+coord_flip()+ylab(NULL)+guides(color="none")+theme(text=element_text(size=16))

ggsave("output/SWIM_figures/Bristol Bay Strahler Order boxplots.jpg", width=16, height=5, units="in", dpi=1000)

summary(BB_mets)

```
#Air temperature plot
Making a plot of longterm temps to put 2019 in context. South Naknek Airport (1994-2024), King Salmon Airport (1942-2024), Iliamna Airport (1948-1971, 1973-2024). 
```{r}
nearest_stations(58.925739, -157.009770, 300)
naknek_air<-get_GSOD(years=1994:2024, station="703335-99999")
naknek_air$Location<-as.factor("Naknek")
kingsalmon_air<-get_GSOD(years=1942:2024, station="703260-25503")
kingsalmon_air$Location<-as.factor("King Salmon")
iliamna_air<-rbind(get_GSOD(years=1948:1971,station="999999-25506"), get_GSOD(years=1973:2024, station="703400-25506"))
iliamna_air$Location<-as.factor("Iliamna")

longterm_BB_air<-rbind(iliamna_air, kingsalmon_air)

BB_air_summer<-longterm_BB_air%>%
  filter(MONTH %in% (6:9))

BB_summary<-BB_air_summer%>%
  group_by(Location, YEAR)%>%
  mutate(mean_temp=mean(TEMP), max_temp=max(TEMP))

ggplot(BB_summary, aes(x=YEAR, y=mean_temp, color=Location))+geom_point()+geom_smooth()+geom_point(data=BB_summary[BB_summary$YEAR==2019,], pch=1, size=5, color="black")+theme_bw()+scale_color_manual(values=c("plum4", "lightsalmon"))+annotate("rect", xmin=2015, xmax=2019, ymin=9, ymax=15, alpha=0.2, fill="red")+annotate("rect", xmin=2001, xmax=2020, ymin=9, ymax=15, alpha=0.2, fill="gray")


```