---
title: "AFS_Data_Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


# load packages
library(googledrive)
library(lubridate)
library(readr)
library(hms)
library(ggmap)
library(sf)
# library(leaflet)
# library(osmdata)
library(broom)
library(caTools)
library(tidyverse)

# install.packages("devtools")
devtools::install_github("yutannihilation/ggsflabel")
library(ggsflabel)

```


notes: fix usgs daily by calculating a mean when only min and max are present.

Bringing in best set of daily data for each data provider and combining into one dataset so I can do some summaries for the AFS presentation.

Bristol Bay

* CIK data ready - saved to final_data folder 16 sites
* ACCS data ready - saved 15 sites
* NPS stream data ready - saved 17 stream sites
* FWS stream data ready - saved 23 stream sites
* UW filter on just TC sites and add in new summer data from Jackie's database - may need to filter out some air temps.

Cook Inlet

* CIK data ready
* ACCS data ready
* USFS data - screen on sites in cook inlet ready
* Deshka project ready
* Kenai project ready
* Anchor-stariski project ready
* USGS data - ready for both ci and bb

Optional
* Thermal regimes data



# Metadata

Bring in the metadata for each set of data and combine so that I can also create some simple maps showing years of data.
Get metadata files off of google drive for both Cook Inlet and Bristol Bay.

```{r read in metadata}
gd.akssf.files <- drive_ls(path = "https://drive.google.com/drive/u/0/folders/1_qtmORSAow1fIxvh116ZP7oKd_PC36L0")

gd.metadata.files <- gd.akssf.files %>% 
  filter(grepl("Metadata", name) & grepl(".csv", name))

folder <- "data_preparation/final_data/Metadata/"

for (i in seq_along(gd.metadata.files$name)) {
  drive_download(as_id(gd.metadata.files$id[i]),
                 path = str_c(folder, gd.metadata.files$name[i]),
                 overwrite = TRUE)
}


local.md.files <- list.files(folder, full.names = TRUE)

local.md.files <- local.md.files[!grepl("Eyak", local.md.files)]

md <- map_df(local.md.files, function(x) read_csv(x, col_types = "ccccccccccccccc") %>%
                      mutate(file_name = basename(x))) %>% 
  mutate(Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

md <- md %>% 
  mutate(SiteID = case_when(is.na(SiteID) ~ Agency_ID,
                            TRUE ~ SiteID))

```
Remove Luca's sites in the Copper River basin. 

```{r remove copper river sites}
luca <- md %>% filter(grepl("Luca", Contact_person), !grepl("Williwaw", SiteID)) %>% pull(SiteID)

md <- md %>% 
  filter(!SiteID %in% luca)
```

Read in HUC8 feature class and identify sites from the Togiak Refuge that are in the Kuskokwim Delta HUC8, outside Bristol Bay.

Create an SF object from the md for intersecting with HUC8s.

```{r create metadata sf}
md_sf <- st_as_sf(md, coords = c("Longitude", "Latitude"), crs = "WGS84")
```

Read in HUC8s and reproject.

```{r add huc8 names to md_sf}
huc8 <- st_read(dsn = "S:/Leslie/GIS/NHD Harmonized/WBD_19_GDB.gdb", layer = "WBDHU8")
st_crs(huc8)
huc8_wgs84 <- st_transform(huc8, crs = "WGS84")

md_sf <- st_join(md_sf, huc8_wgs84)

```

Remove Kuskokwim Delta sites from md and md_sf.

```{r remove kuskokwim sites}
kusko <- md_sf %>% filter(Name == "Kuskokwim Delta") %>% pull(SiteID)

md <- md %>% 
  filter(!SiteID %in% kusko)

md_sf <- md_sf %>% 
  filter(!SiteID %in% kusko)

```

Add in a region - Bristol Bay or Cook Inlet. The file names can't be used since the USGS data are from both BB and CI.

```{r add region to md}
ci_hucs <- st_drop_geometry(md_sf) %>% 
  filter(HUC8 < 19030000) %>% 
  distinct(HUC8) %>% 
  pull(HUC8)

md_sf <- md_sf %>% 
  mutate(Region = case_when(HUC8 %in% ci_hucs ~ "Cook Inlet",
                            TRUE ~ "Bristol Bay"))

md <- left_join(md, st_drop_geometry(md_sf) %>% distinct(SiteID, HUC8, Name, Region))
```


# Data

Get daily files off of google drive for both Cook Inlet and Bristol Bay.

```{r read in daily data}
gd.daily.files <- gd.akssf.files %>% 
  filter(grepl("Daily_Data", name) & grepl(".csv", name))

folder <- "data_preparation/final_data/Daily_Data/"

for (i in seq_along(gd.daily.files$name)) {
  drive_download(as_id(gd.daily.files$id[i]),
                 path = str_c(folder, gd.daily.files$name[i]),
                 overwrite = TRUE)
}


local.daily.files <- list.files(folder, full.names = TRUE)

daily.dat <- map_df(local.daily.files, function(x) read_csv(x, col_types = "cDnnn") %>%
                      mutate(file_name = basename(x)))

```

Remove Luca's sites in the Copper River basin and Togiak Refuge sites from Kuskokwim. 

```{r remove copper and kuskokwim sites}
daily.dat <- daily.dat %>% 
  filter(!SiteID %in% c(luca, kusko))

#check
daily.dat %>% 
  distinct(SiteID, file_name) %>% 
  count(file_name) %>% 
  rename(daily_n = n) %>% bind_cols(md %>% 
  distinct(SiteID, file_name) %>% 
  count(file_name))
```

Filter to only include site years with 80% of days in the summertime (6-8).

```{r create daily.sum}

daily.dat %>% 
  filter(is.na(meanDT))

daily.sum <- daily.dat %>%
  filter(!is.na(meanDT)) %>% 
  group_by(SiteID, Year = year(sampleDate)) %>% 
  filter(month(sampleDate) %in% 6:8) %>% 
  mutate(summer_ct = n()) %>% 
  filter(summer_ct > 0.8 * 92) %>% 
  left_join(md %>% select(SiteID, Name, Region)) %>% 
  ungroup()

nrow(daily.dat %>% distinct(SiteID, Year = year(sampleDate)))
nrow(daily.sum %>% distinct(SiteID, Year = year(sampleDate)))

#397 site years lost because incomplete.
1897-1500
```

Calculate thermal regime metrics for summer data only - June through August.

* Maximum mean daily temperature 
* MWAT Max of 7-day rolling average of mean daily temp
* Timing of MWAT - low priority


```{r mets from daily mean}
max_meanDT <- daily.sum %>%
  filter(!is.na(meanDT)) %>% 
  group_by(SiteID, Year = year(sampleDate)) %>% 
  summarize(max_DAT = max(meanDT, na.rm = TRUE))
  
MWAT <- daily.sum %>%
  filter(!is.na(meanDT)) %>% 
  group_by(SiteID, Year = year(sampleDate)) %>% 
  summarize(MWAT = max(runmean(meanDT, k = 7, endrule = "NA", align = "center"), na.rm=TRUE))


mets <- left_join(max_meanDT, MWAT)
mets <- left_join(mets, md %>% select(SiteID, Name, Region))
```


Old code chunk when we had min mean and max for all sites. UW data is mean only so just calculating metrics for plotting.

```{r temperature metrics, eval = FALSE}
source('W:/Github/SWSHP-Bristol-Bay-Thermal-Diversity/Temperature Descriptor Function - daily inputs.R')

#count missing max and min for usgs sites - two sites missing max and min altogether, just remove
daily.dat %>% 
  group_by(SiteID) %>% 
  summarize(mxct = sum(!is.na(maxDT)),
            mnct = sum(!is.na(minDT)),
            meanct = sum(!is.na(meanDT)))

nrow(daily.dat)
nrow(daily.dat %>% na.omit(.))

mets.input <- daily.dat %>%
  na.omit(.) %>% 
  group_by(SiteID, Year = year(sampleDate), month = month(sampleDate)) %>% 
  filter(month %in% 6:8) %>% 
  mutate(mon_total = days_in_month(month),
         mon_ct = n()) %>% 
  filter(mon_ct > 0.8 * mon_total)


# daily.sum <- daily.dat %>% 
#   rename(site_name = SiteID, date = sampleDate, mean = meanDT, min = minDT, max = maxDT) %>% 
#   tempscreen()

#modified so no longer saving to excel file, I just need the data frame output.
mets <- daily.sum %>% 
  rename(site_name = SiteID, date = sampleDate, mean = meanDT, min = minDT, max = maxDT) %>% 
  mutate(site.year = paste(site_name,year(date),sep=".")) %>% 
  as.data.frame() %>% 
  tempmetrics(., "output/metrics_2021-03-11")

mets <- mets %>% 
  mutate(Site = gsub("\\..*","", site.year),
         Year = gsub("^.*\\.","", site.year))

```


# HUC8 Summaries

SF objects for md and HUC8s already created above. And HUC8 names to daily data by joining to md.

Data summary table by HUC8. Add back to the HUC8 SF object for mapping.

* count of sites in a HUC8, 
* count of years in a HUC8, 
* number of long-term sites in a HUC8, > 9 years of data
* number of sites with at least 3 years of data in a HUC8


```{r data availability by huc8}

# use daily.sum because that is filtered on sites with 80% of data for summer months.

huc8.siteCt <- daily.sum %>% 
  distinct(Region, Name, SiteID) %>% 
  count(Region, Name) %>% 
  rename(Site_Count = n)

huc8.yearCt <- daily.sum %>% 
  distinct(Region, Name, SiteID, Year = year(sampleDate)) %>% 
  count(Region, Name) %>% 
  rename(Year_Count = n)

huc8.LTsiteCt <- daily.sum %>% 
  distinct(Region, Name, SiteID, Year = year(sampleDate)) %>% 
  count(Region, Name, SiteID) %>%
  filter(n > 9) %>% 
  count(Region, Name) %>% 
  rename(LT_Site_Count = n)

huc8.3yrsiteCt <- daily.sum %>% 
  distinct(Region, Name, SiteID, Year = year(sampleDate)) %>% 
  count(Region, Name, SiteID) %>%
  filter(n > 2) %>% 
  count(Region, Name) %>% 
  rename(Yr3_Site_Count = n)

```

Create HUC8 SF for just COok Inlet and Bristol Bay and add the summaries to the object for mapping.

```{r filter HUC8s to CI and BB and add summaries}
#get vector of HUC8 names for just ci and bb -- study area (sa)
huc8_names_sa <- st_drop_geometry(md_sf) %>% distinct(Name) %>% pull(Name)

huc8_sa <- huc8_wgs84 %>% 
  filter(Name %in% c(huc8_names_sa)) %>% 
  left_join(huc8.siteCt) %>% 
  left_join(huc8.yearCt) %>% 
  left_join(huc8.LTsiteCt) %>% 
  left_join(huc8.3yrsiteCt)

```


```{r create cities sf}
cities <- st_as_sf(data.frame(place = c("Dillingham", "Anchorage"),
                              longitude = c(-158.508665, -149.9),
                              latitude = c(59.046751, 61.216667)), coords = c("longitude", "latitude"), 
                   crs = "WGS84")

```
 

# Figures


```{r boxplot of mean july temperatures}
mon.summ <- daily.sum %>% 
  filter(!is.na(meanDT)) %>% 
  group_by(Region, Name, SiteID, month = month(sampleDate), year = year(sampleDate)) %>% 
  summarize(mon_mn = mean(meanDT))  

site.order <- mon.summ %>%
  filter(month == 7) %>% 
  group_by(Name) %>% 
  summarize(mean_huc = mean(mon_mn)) %>% 
  arrange(desc(mean_huc)) %>% 
  pull(Name)

mon.summ %>% 
  filter(month == 7) %>% 
  mutate(Namef = factor(Name, levels = site.order)) %>%
  ggplot() +
  geom_boxplot(aes(y = Namef, color = Region, x = mon_mn)) +
  geom_text(inherit.aes = FALSE, data = . %>% group_by(Namef) %>% count(), 
            aes(label = paste0("(", n, ")"), y = Namef, x = -0.2), size = 3) +
  theme_minimal() +
  theme(legend.position = "bottom", axis.title.y = element_blank()) +
  labs(x = "Temperature (째C)",
       title = "Mean July Temperatures by Watershed",
       subtitle = "Bristol Bay and Cook Inlet")
```


```{r daily time series}

hucs9 <- huc8.siteCt %>% arrange(desc(Site_Count)) %>% filter(Site_Count > 9) %>% pull(Name)

daily.sum %>% 
  mutate(year = year(sampleDate)) %>% 
  filter(Name %in% hucs9) %>% 
  distinct(year) %>% 
  arrange(year)

daily.sum %>% 
  mutate(year = year(sampleDate)) %>% 
  filter(Name %in% hucs9) %>% 
  distinct(SiteID, year) %>% 
  count(SiteID) %>% 
  arrange(n)

daily.sum %>% 
  filter(Name %in% hucs9) %>% 
  mutate(year = year(sampleDate),
         mo_day = format(sampleDate, "%m-%d"),
         site_year = paste0(SiteID, year),
         is_19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0)) %>%
  filter(month(sampleDate) %in% 6:9) %>% 
  ggplot() +
  geom_line(data = . %>% filter(is_19 == 0), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year), color = "grey") +
  geom_line(data = . %>% filter(is_19 == 1), 
            aes(x = as.Date(mo_day, format = "%m-%d"), y = meanDT, group = site_year, color = "red")) +
  geom_abline(aes(intercept = 20, slope = 0), linetype = 2) +
  facet_wrap(~Name) +
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  labs(title = "Stream Temperatures for Cook Inlet and Bristol Bay Watersheds",
       subtitle = "299 sites with one to 23 years of data from 1975-2020 \n2019 data in red",
       y = "Mean Daily Temperature (째C)")

ggsave("output/AFS/Daily Temps for Nine HUCs.jpg", units = "in", width = 10, height = 7.5)
```

Dotplot of metrics

```{r dotplot of MWAT}
mwat.order <- mets %>%
  group_by(SiteID) %>% 
  summarize(mn_mwat = mean(MWAT)) %>% 
  arrange(mn_mwat) %>% 
  pull(SiteID)

mets %>% 
  filter(Name == "Lower Susitna River")

names.filter <- c("Lake Clark", "Togiak", "Lower Susitna River", "Anchorage")

mets %>%
  mutate(SiteIDf = factor(SiteID, levels = mwat.order)) %>% 
  # filter(Name %in% names.filter) %>% 
  ggplot() +
  geom_point(data = . %>% filter(Year != 2019), aes(x = MWAT, y = SiteIDf, color = Region)) +
  geom_point(data = . %>% filter(Year == 2019), aes(x = MWAT, y = SiteIDf), color = "black") +
  geom_vline(aes(xintercept = 13), linetype = 2) +
  geom_vline(aes(xintercept = 18), linetype = 2) +
  geom_vline(aes(xintercept = 20), linetype = 2) +
  facet_wrap(~Name, scales = "free") +
  theme_bw() +
  theme(axis.text.y = element_blank()) 
```

```{r boxplot of MWAT 9 HUCs}
huc.mwat.order <- mets %>%
  filter(Name %in% hucs9) %>% 
  group_by(Name) %>% 
  summarize(mean_huc = mean(MWAT)) %>% 
  arrange(desc(mean_huc)) %>% 
  pull(Name)

mets %>% 
  mutate(Namef = factor(Name, levels = huc.mwat.order)) %>%
  filter(Year != 2019, Name %in% hucs9) %>% 
  ggplot() +
  geom_boxplot(aes(y = Namef, color = Region, x = MWAT)) +
  geom_text(inherit.aes = FALSE, data = . %>% group_by(Namef) %>% count(), 
            aes(label = paste0("(", n, ")"), y = Namef, x = -0.2), size = 3) +
  theme_minimal() +
  theme(legend.position = "bottom", axis.title.y = element_blank()) +
  labs(x = "Temperature (째C)",
       title = "MWAT by Watershed",
       subtitle = "Bristol Bay and Cook Inlet")
```

CIK_28 = deshka, why only 5 years of data? -- lots of incomplete years for evaluating summer thermal regimes.

```{r check Deshka data}
daily.dat %>% 
  filter(SiteID == "CIK_28") %>%
  group_by(year(sampleDate)) %>% 
  summarize(min(sampleDate),
             max(sampleDate))
daily.sum %>% 
  filter(SiteID == "CIK_28") %>% 
  distinct(year(sampleDate))

```


```{r boxplot of MWAT LT Sites}

ltSites <- daily.sum %>% 
  left_join(md %>% distinct(SiteID, Waterbody_name)) %>% 
  distinct(Region, Name, SiteID, Waterbody_name, Year = year(sampleDate)) %>% 
  count(Region, Name, Waterbody_name, SiteID) %>%
  filter(n > 9) %>% 
  pull(SiteID)

lt.mwat.order <- mets %>%
  filter(SiteID %in% ltSites) %>% 
  left_join(md %>% select(SiteID, Waterbody_name)) %>% 
  mutate(Waterbody_name = case_when(SiteID == "Aleknagik Bear Creek" ~ "Bear Cr",
                                    SiteID == "15261000" ~ "Cooper Cr",
                                    SiteID == "15283700" ~ "Moose Cr nr Palmer",
                                    SiteID == "15300250" ~ "Upper Talarik Cr",
                                    SiteID == "15239070" ~ "Bradley R",
                                    SiteID == "15302200" ~ "Koktuli R",
                                    SiteID == "15258000" ~ "Kenai R at Cooper Landing",
                                    SiteID == "15302250" ~ "NF Koktuli R",
                                    SiteID == "15266300" ~ "Kenai R at Soldotna",
                                    SiteID == "15300100" ~ "Bear Cr nr Iliamna",
                                    SiteID == "PULR" ~ "Pungokepuk-Downstream",
                                    SiteID == "PULO" ~ "Pungokepuk-Lake Outlet",
                                    TRUE ~ Waterbody_name)) %>% 
  group_by(Waterbody_name, SiteID) %>% 
  summarize(mwat.mn = mean(MWAT),
            n = n()) %>% 
  arrange(desc(mwat.mn)) #%>% 
  pull(Waterbody_name)

# lt.mwat.order <- mets %>%
#   filter(SiteID %in% ltSites) %>% 
#   left_join(md %>% select(SiteID, Waterbody_name)) %>% 
#   group_by(Waterbody_name, SiteID) %>% 
#   summarize(mwat.mn = mean(MWAT),
#             n = n()) %>% 
#   arrange(desc(mwat.mn)) %>% 
#   pull(SiteID)

mets %>% 
  filter(SiteID %in% ltSites) %>% 
  left_join(lt.mwat.order %>% select(SiteID, Waterbody_name)) %>% 
  mutate(wbf = factor(Waterbody_name, levels = lt.mwat.order %>% pull(Waterbody_name))) %>%
  ggplot() +
  geom_boxplot(data = . %>% filter(Year != 2019), aes(y = wbf, color = Region, x = MWAT)) +
  # geom_text(inherit.aes = FALSE, data = . %>% group_by(SiteIDf) %>% count(), 
  #           aes(label = paste0("(", n, ")"), y = SiteIDf, x = -0.2), size = 3) +
  geom_point(data = . %>% filter(Year == 2019), aes(y = wbf, x = MWAT, fill = "2019"), shape = 15, size = 1.5) +
  theme_minimal() +
  theme(legend.position = "bottom", axis.title.y = element_blank()) +
  labs(x = "Temperature (째C)",
       title = paste("Maximum Weekly Average Temperature for \nSites with", "\u2265 10", "Years of Data"),
       fill = "")

ggsave("output/AFS/MWAT for long-term sites.jpg", units = "in", width = 10, height = 7.5)
```



```{r monthly data frame}

month.dat <- daily.dat %>% 
  mutate(daysMonth = days_in_month(sampleDate),
         month = month(sampleDate),
         year = year(sampleDate)) %>% 
  group_by(SiteID, year, month, daysMonth) %>%
  summarize(meanMon = mean(meanDT),
            monCt = n()) %>% 
  filter(monCt > (0.9 * daysMonth))
```



```{r trends in June temps}
junSites <- month.dat %>% 
  ungroup() %>% 
  filter(month %in% 6) %>% 
  distinct(SiteID, year) %>% 
  count(SiteID) %>% 
  filter(n > 10) %>% 
  pull(SiteID)

#June
left_join(month.dat, md %>% select(SiteID, Waterbody_name, Name, Region)) %>% 
  filter(SiteID %in% junSites, month %in% 6) %>% 
  ggplot(aes(x = year, y = meanMon, color = Waterbody_name)) + 
  geom_point() +
  # geom_line() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(2000, 2020)) +
  facet_wrap(~Region) +
  theme_minimal() +
  theme(legend.position = "none")


```

```{r linear reg. of monthly means}
ltMonths <- month.dat %>% 
  ungroup() %>% 
  distinct(SiteID, year, month) %>% 
  count(SiteID, month) %>% 
  filter(n > 10)

monthlyLMs <- month.dat %>% 
  filter(!year == 2019) %>% 
  right_join(ltMonths) %>% 
  nest(data = c(-SiteID, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(meanMon ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied) 

sig_trends <- monthlyLMs %>% 
  filter(p.value < 0.05, term == "year") %>% 
  select(month)

left_join(monthlyLMs, md %>% select(SiteID, Waterbody_name, Name, Region)) %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate, color = Waterbody_name)) +
  geom_point() +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0))  +
  facet_wrap(~Region) +
  theme(legend.position = "bottom")


```



```{r fig.height = 11, fig.width = 7}

bb.mon <- bbdat %>% 
  mutate(mo_days = days_in_month(sampleDate)) %>% 
  group_by(Name, SiteID, mo_days, month = month(sampleDate), year = year(sampleDate)) %>% 
  summarize(mean = mean(meanDT),
            mo_ct = n()) %>%
  filter(mo_ct > 0.9 * mo_days)

bb.july.n5 <- bb.mon %>% 
  filter(month == 7) %>% 
  group_by(SiteID) %>% 
  mutate(ct = n()) %>% 
  filter(ct > 4)

order19 <- bb.july.n5 %>% group_by(SiteID) %>% filter(month == 7) %>% summarize(mean7 = mean(mean)) %>% arrange(desc(mean7)) %>% pull(SiteID)

bb.july.n5 <- bb.july.n5 %>% 
  mutate(SiteIDf = factor(SiteID, levels = order19))


ggplot() +
  geom_boxplot(data = bb.july.n5 %>% filter(month == 7),aes(x = mean, y = SiteIDf, color = Name))  +
  geom_point(data = bb.july.n5 %>% filter(year == 2019, month ==7), aes(x = mean, y = SiteIDf), color = "red") +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  labs(x = "Mean July Temperature", title = "Mean July Temperatures for 71 Streams in Bristol Bay with \nFive or More Years of Data",
       subtitle = "Red dots indicate mean July temperatures for sites with data from 2019")

```





# Mapping


## Data Availability Maps

```{r number of sites in each huc8}
huc8_reg <- huc8_sa %>% 
  group_by(Region) %>% 
  summarize()

huc8_sa %>% arrange(Site_Count) %>% pull(Site_Count) 
huc8_sa %>% st_drop_geometry() %>% arrange(Site_Count) %>% summarize(sum(Site_Count, na.rm = TRUE))

site_yrCt <- daily.sum %>% 
  distinct(SiteID, Year) %>% 
  count(SiteID) %>% 
  rename(yrct = n)

md_sf <- left_join(md_sf, site_yrCt)

huc8_sa %>% 
  mutate(cutn = cut(Site_Count, breaks = c(0, 5, 10, 30, 130), 
                    labels = c("1-5", "6-10", "11-30", "31-125"))) %>%
  ggplot() +
  geom_sf(aes(fill = cutn), color = "gray") + 
  geom_sf(data = huc8_reg, color = "black", fill = NA, lwd = 1) +
  geom_sf(data = md_sf, size = 1.2) +
  # geom_sf(data = md_sf, aes(size = yrct), shape = 21, fill = NA) +
  # geom_sf(data = md_sf, color = "black", shape = 21, fill = NA, size = 1.5) +
  geom_sf_label_repel(data = huc8_reg, aes(label = Region), nudge_x = -1.5, nudge_y = 2) +
  annotate(geom = "text", x = -160, y = 63, label = "N = 341 Sites", size = 4) +
  scale_fill_brewer(type = "seq", palette = 3) +
  # theme_minimal() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  labs(fill = "Number of Sites: ")

ggsave("output/AFS/Sites by HUC.jpeg", units = "in", height = 7.5, width = 10)
```


```{r number of years in each huc8}
huc8_sa %>% arrange(Year_Count) %>% pull(Year_Count) 
huc8_sa %>% st_drop_geometry() %>% arrange(Year_Count) %>% summarize(sum(Year_Count, na.rm = TRUE))

huc8_sa %>% 
  mutate(cutn = cut(Year_Count, breaks = c(0, 10, 50, 100, 320), 
                    labels = c("4-10", "11-50", "51-100", "101-316"))) %>%
  ggplot() +
  geom_sf(aes(fill = cutn), color = "gray") + 
  geom_sf(data = huc8_reg, color = "black", fill = NA, lwd = 1) +
  geom_sf(data = md_sf, size = 1.2) +
  geom_sf_label_repel(data = huc8_reg, aes(label = Region), nudge_x = -1.5, nudge_y = 2) +
  annotate(geom = "text", x = -160, y = 63, label = "N = 1500 Years", size = 4) +
  scale_fill_brewer(type = "seq", palette = 3) +
  # theme_minimal() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  labs(fill = "Number of Years: ")

ggsave("output/AFS/Years by HUC.jpeg", units = "in", height = 7.5, width = 10)
```

```{r Location of long-term sites}



huc8_sa %>% arrange(LT_Site_Count) %>% pull(LT_Site_Count) 
huc8_sa %>% st_drop_geometry() %>% arrange(LT_Site_Count) %>% summarize(sum(LT_Site_Count, na.rm = TRUE))

huc8_sa %>% 
  mutate(cutn = cut(LT_Site_Count, breaks = c(0, 3, 7), 
                    labels = c("1-3", "4-7"))) %>%
  ggplot() +
  geom_sf(aes(fill = cutn), color = "gray") + 
  geom_sf(data = huc8_reg, color = "black", fill = NA, lwd = 1) +
  geom_sf(data = md_sf %>% filter(SiteID %in% ltSites), size = 1.2) +
  geom_sf_label_repel(data = huc8_reg, aes(label = Region), nudge_x = -1.5, nudge_y = 2) +
  annotate(geom = "text", x = -159, y = 63, label = "N = 36 Long-Term Sites", size = 4) +
  scale_fill_brewer(type = "seq", palette = 3) +
  # theme_minimal() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  labs(fill = paste("Number of Sites with", "\u2265 10", "Years of Data: "))

ggsave("output/AFS/LT Sites by HUC.jpeg", units = "in", height = 7.5, width = 10)
```


```{r Location of sites with 3 years of data}

huc8_sa %>% arrange(Yr3_Site_Count) %>% pull(Yr3_Site_Count) 
huc8_sa %>% st_drop_geometry() %>% arrange(Yr3_Site_Count) %>% summarize(sum(Yr3_Site_Count, na.rm = TRUE))

yr3Sites <- daily.sum %>% 
  distinct(Region, Name, SiteID, Year = year(sampleDate)) %>% 
  count(Region, Name, SiteID) %>%
  filter(n > 2) %>% 
  pull(SiteID)

huc8_sa %>% 
  mutate(cutn = cut(Yr3_Site_Count, breaks = c(0, 5, 20, 60), 
                    labels = c("1-5", "6-20", "21-59"))) %>%
  ggplot() +
  geom_sf(aes(fill = cutn), color = "gray") + 
  geom_sf(data = huc8_reg, color = "black", fill = NA, lwd = 1) +
  geom_sf(data = md_sf, size = 1.2, color = "red") +
  geom_sf(data = md_sf %>% filter(SiteID %in% yr3Sites), size = 1.2) +
  geom_sf_label_repel(data = huc8_reg, aes(label = Region), nudge_x = -1.5, nudge_y = 2) +
  annotate(geom = "text", x = -160, y = 63, label = "N = 214 Sites", size = 4) +
  scale_fill_brewer(type = "seq", palette = 3) +
  # theme_minimal() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  labs(fill = paste("Number of Sites with", "\u2265 3", "Years of Data: "))

ggsave("output/AFS/Sites with 3 yrs of data by HUC.jpeg", units = "in", height = 7.5, width = 10)
```




```{r site count map old}

huc8_sa %>% names

ggplot() +
  geom_sf(data = huc8_sa, aes(fill = Site_Count, color = Region)) +
  geom_sf(data = md_sf, color = "white", shape = 21, fill = NA, size = 2) +
  # geom_sf(data = md_sf, color = "black", size = 0.5) +
  geom_sf_label_repel(data = cities, aes(label = place), nudge_x = -3, size = 3, nudge_y = 1.5) +
  scale_fill_viridis_b() +
  # theme_minimal() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  labs(fill = "Number of Sites in each HUC8")

ggsave("output/AFS/Sites by HUC.jpeg", units = "in", height = 7.5, width = 10)
```





```{r}
register_google(key = "", write = TRUE) #saved locally outside github repo folders
```


```{r basic map}

location <- c(min(md$Longitude), min(md$Latitude), max(md$Longitude), max(md$Latitude))

map <- get_map(location, maptype = "terrain", source = "google")

ggmap(map)

md %>% 
  select(Latitude, Longitude) %>% 
  qmplot(Longitude, Latitude, data = ., source = "google", maptype = "terrain",
         legend = "bottom", zoom = 6)

```

