---
title: "data_USFS_Chugach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(stringr)

```


# USFS

Export his site names for Priscilla or use his metadata workbook -- it's a little confusing. Check if it matches the actual data files.

```{r}
fs_files <- list.files("S:\\Stream Temperature Data\\Luca Adelfio\\ChugachNF_wtdat", full.names = TRUE)

#remove metadata sheet
fs_files <- fs_files[!grepl("siteinfo", fs_files)]

#remove air temperature data sheets
fs_files <- fs_files[!grepl("AirT", fs_files)]

#save names so Priscilla can look up AKOATS ID
fs_site_names <- sapply(fs_files, function(x) str_match(x, "\\/\\s*(.*?)\\s*\\.csv")[2], simplify = TRUE)
as.data.frame(fs_site_names) %>% write_csv(path = "output/usfs_site_names.csv")

```

# Combine data across csv files

```{r}

read_csv_and_name <- function(csv_file_path) {
  sheet_name <- str_match(csv_file_path, "\\/\\s*(.*?)\\s*\\.csv")[2]
  dat <- read_csv(csv_file_path) %>% 
      mutate(SiteID = sheet_name)
}

fsdat <- fs_files %>% 
  map_df(function(x) read_csv_and_name(x)) 

```

Some problems with dates - some are 2 digit and some are 4 digit. In as.Date function, tryFormats won't fix this problem, which is unfortunate. I'll need to get number of digits using regexp and then manually tell it what format it is in.

```{r}
fsdat %>% mutate(year1 = sub(".*/.*/", "", Date)) %>% distinct(year1)

fsdat <- fsdat %>% 
  mutate(year1 = sub(".*/.*/", "", Date),
         sampleDate = case_when(nchar(year1) == 4 ~ as.Date(Date, format = "%m/%d/%Y"),
                                TRUE ~ as.Date(Date, format = "%m/%d/%y"))) %>% 
  rename(sampleTime = Time,
         Temperature = Temp)

fsdat %>% 
  group_by(SiteID) %>% 
  summarize(min(sampleDate), max(sampleDate))

fsdat %>% 
  distinct(year(sampleDate))

```

Quick plots of data to make sure they read in ok. Start by summarizing daily means because quicker to plot. Looks like some bad winter temps well below zero that could be clipped later. No obvious air temps in summer as everything is < = 20 or so.

```{r}
fsdat %>% 
  group_by(SiteID, sampleDate) %>% 
  summarize(meant = mean(Temperature)) %>% 
  ggplot(aes(x = sampleDate, y = meant)) +
  geom_line() +
  facet_wrap(~SiteID)

```
Wrong! The sub-daily temps show more erros, definitely some air temps that need to be removed. This will be a good dataset for testing scripts, although it will probably need cleaning sooner rather than later for AKSSF.

```{r}
fsdat %>% 
  mutate(dt = as.POSIXct(paste(sampleDate, sampleTime, sep = " "), format = "%Y-%m-%d %H:%M")) %>% 
  ggplot(aes(x = dt, y = Temperature)) +
  geom_line() +
  facet_wrap(~SiteID)

```
Save as csv for copying over to the drive.

```{r}

```

