---
title: "data_lakes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))


library(readxl)
library(tidyverse)
# library(dygraphs)
# library(rgl)
library(lubridate)
library(gridExtra)
```

Note that NPS and FWS lake data are included in their respective scripts. This is a separate script for UW lake data.

# UW lake data

Read in UW data frame - this is the original file with depths only to 5 m.

```{r read in file with data to 5m}
uw_lak <- read_excel("S:\\Stream Temperature Data\\Jackie Carter UW\\Lake Data\\Lake_Temps_for Becky.xlsx")

summary(uw_lak)

uw_lak %>% 
  distinct(Lake, YearSampled, DateSampled) %>% 
  group_by(Lake, YearSampled) %>%
  summarize(count = n(),
            startDate = min(DateSampled),
            endDate = max(DateSampled))

uw_lak %>% 
  group_by(Lake) %>% 
  summarize(startYear = min(YearSampled),
            endYear = max(YearSampled)) %>% 
  write_csv("output/uw_lakes.csv")

uw_lak %>% 
  group_by(Lake) %>% 
  summarize(min(Depth),
            max(Depth))

uw_lak %>% count(Notes)
uw_lak %>% summary(Depth)  
```

```{r plot of data to 5m}
uw_lak %>% 
  filter(Lake == "Aleknagik") %>% 
  ggplot(aes(x = DateSampled, y = Depth, color = Temperature, group = DateSampled)) +
  geom_point() +
  facet_wrap(~StationName)
```

New lake data files from Jackie that include the entire depth profiles. Series of 5 files that all start "forRS".

```{r read in complete depth arrays}

uw_lake_files <- list.files("S:\\Stream Temperature Data\\Jackie Carter UW\\Lake Data\\", full.names = TRUE)
uw_lake_files <- uw_lake_files[grepl("forRS", uw_lake_files)]

uw_lkdat <- map_df(uw_lake_files, function(x) read_excel(x))
```

Clean up Lake capitalization and flag strange temperatures at Aleknagik Bear Bay Pt in 2006 - very warm.

```{r}

uw_lkdat %>% distinct(Lake)
uw_lkdat %>% filter(Temperature > 22)

uw_lkdat <- uw_lkdat %>% 
  mutate(Lake = case_when(Lake == "aleknagik" ~ "Aleknagik", 
                          TRUE ~ Lake),
         UseData = case_when(Temperature > 22 & StationName == "Aleknagik Bear Bay Pt" ~ 0, 
                             TRUE ~ 1),
         DateSampled = as.Date(DateSampled))


```

What are the frequencies of these depth profiles? It looks like some lakes may be weekly over the summer, whereas others may just have one profile.

```{r profile counts by site and year}
uw_lkdat %>% 
  distinct(Lake, StationName, DateSampled) %>% 
  mutate(year = year(DateSampled)) %>% 
  count(Lake, StationName, year) %>% 
  ggplot(aes(x = year, y = StationName, size = n, color = Lake)) +
  facet_wrap(~Lake, scales = "free_y") +
  geom_point() +
  theme_bw()

```

Rolling pdf of lake plots.

```{r eval = FALSE}
uw_lake_sites <- uw_lkdat %>% distinct(StationName) %>% pull(StationName)

pdf("output/UW Lake Plots.pdf")
for(i in uw_lake_sites) {
  dat <- uw_lkdat %>% 
    filter(StationName == i) %>% 
    mutate(year = year(DateSampled)) 

  p <- ggplot(data = dat, aes(x = DateSampled, y = Depth, color = Temperature)) +
    geom_point() +
    scale_color_gradient2(limits = c(0, 25), midpoint = 10, high = scales::muted("red"),
                          low = scales::muted("blue")) +
    scale_y_reverse() +
    facet_wrap(~year, scales = "free_x") +
    theme_bw() +
    labs(title = dat[1, "Lake"], subtitle = dat[1, "StationName"])
  print(p)
}
dev.off()
```

Interpolate temperatures by depth and date for each site and year to create heatmaps.
Follow guidance on this [website for depth-time heatmaps] (https://fishandwhistle.net/post/2019/depth-time-heatmaps/).

```{r heatmap of full depth profiles}

#first step is interpolating depths, but the depth data are mostly continuous
estimate_temp_by_date <- function(target_date, target_depth) {
  data_for_date <- uw_lkdat %>% 
    filter(DateSampled == target_date) %>%
    arrange(Depth)
  
  # approx() is one way to do a linear interpolation
  approx(data_for_date$Depth, data_for_date$Temperature, xout = target_depth)$y
}

temp_interp_depth <- crossing(
  # the same dates as sonde_tbl_1993
  tibble(date = unique(sonde_tbl_1993$date)),
  # depths can now be any value
  tibble(depth = seq(1, 20, length.out = 100))
) %>%
  group_by(date) %>%
  mutate(temp = estimate_temp_by_date(date[1], depth))

# create a function that will, given a depth, estimate the temp on any given day
estimate_temp_by_depth <- function(target_depth, target_date) {
  data_for_depth <- temp_interp_depth %>% 
    filter(Depth == target_depth) %>%
    arrange(DateSampled)
  approx(data_for_depth$DateSampled, data_for_depth$Temperature, xout = target_date)$y
}

temp_raster <- crossing(
  # dates can now be any value
  tibble(date = seq(ymd("1993-05-13"), ymd("1993-10-06"), by = 1)),
  # depths must be the same as in temp_interp_depth
  tibble(depth = unique(temp_interp_depth$depth))
) %>%
  group_by(depth) %>%
  mutate(temp = estimate_temp_by_depth(depth[1], date))
```

Try the interpolation function for dates on one lake station and year.

```{r}
alekbearpt_2015 <- uw_lkdat %>% 
  filter(StationName == "Aleknagik Bear Bay Pt", year(DateSampled) == 2015)

estimate_temp_by_date <- function(target_date, target_depth) {
  data_for_date <- alekbearpt_2015 %>% 
    filter(DateSampled == target_date) %>%
    arrange(Depth)
  
  # approx() is one way to do a linear interpolation
  approx(data_for_date$Depth, data_for_date$Temperature, xout = target_depth)$y
}

temp_interp_depth <- crossing(
  # the same dates as sonde_tbl_1993
  tibble(DateSampled = unique(alekbearpt_2015$DateSampled)),
  # depths can now be any value
  tibble(Depth = seq(1, 45, length.out = 100))
) %>%
  group_by(DateSampled) %>%
  mutate(Temperature = estimate_temp_by_date(DateSampled[1], Depth))


p1 <- ggplot(data = temp_interp_depth, aes(x = DateSampled, y = Depth, color = Temperature)) +
    geom_point() +
    scale_color_gradient2(limits = c(0, 25), midpoint = 10, high = scales::muted("red"),
                          low = scales::muted("blue")) +
    scale_y_reverse() +
    theme_bw()

estimate_temp_by_depth <- function(target_depth, target_date) {
  data_for_depth <- temp_interp_depth %>% 
    filter(Depth == target_depth) %>%
    arrange(DateSampled)
  approx(data_for_depth$DateSampled, data_for_depth$Temperature, xout = target_date)$y
}


temp_raster <- crossing(
  # dates can now be any value
  tibble(DateSampled = seq(ymd("2015-06-05"), ymd("2015-09-10"), by = 1)),
  # depths must be the same as in temp_interp_depth
  tibble(Depth = unique(temp_interp_depth$Depth))
) %>%
  group_by(Depth) %>%
  mutate(Temperature = estimate_temp_by_depth(Depth[1], DateSampled))

p2 <- ggplot(temp_raster, aes(DateSampled, Depth, fill = Temperature)) +
  geom_raster() +
  scale_y_reverse() +
  scale_fill_gradient2(
    midpoint = 10, 
    high = scales::muted("red"), 
    low = scales::muted("blue")
  ) +
  coord_cartesian(expand = FALSE)

grid.arrange(p1, p2, ncol = 2)

```



```{r}
with(uw_lkdat %>% filter(StationName == "Aleknagik Agulowak"), plot3d(DateSampled, Depth, Temperature, type = "l", col = as.numeric(Temperature)))
```



