---
title: "Cook Inletkeeper Water Temps"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Rebecca Shaftel (rsshaftel@alaska.edu).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))



# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
library(dataRetrieval)
library(googledrive)
```

Sue used stream names for her site IDs in Bristol Bay. Add "cik_" as a prefix to each in both the data and the metadata so we don't have duplicates later (e.g. Bear Creek x n).

# Metadata

AKOATS metadata

```{r}
akoats.meta <- read_excel("data/AKOATS_DATA_2020_Working.xlsx", sheet = "AKOATS_COMPLETE", col_types = "text") %>%
  filter(Sample_interval == "continuous") %>% 
  select(seq_id,Agency_ID,Contact_person,SourceName,Contact_email,Contact_telephone,Latitude,Longitude,Sensor_accuracy,Waterbody_name,Waterbody_type) %>%
  mutate(seq_id = as.numeric(seq_id),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

akoats.meta
```

CIK metadata

Sue didn't provide lat/longs in her metadata file so need to try and match names in akoats. Sue provided lat/long for Ben Courtney Creek, Dan provided lat/long for Roadhouse creek.

```{r}

cik.meta <- akoats.meta %>%
  filter(SourceName == "CIK") %>% 
  select(seq_id, Agency_ID, SourceName, Contact_person, Waterbody_name, Waterbody_type, Latitude, Longitude)

#remove sites in cook inlet.
cik.meta <- cik.meta[!grepl("CIK", cik.meta$Agency_ID),]

cik.meta <- bind_rows(cik.meta,
                      data.frame(Agency_ID = "Ben Courtney Creek",
                                 SourceName = "CIK",
                                 Contact_person = "Sue Mauger",
                                 Latitude = 59.272417,
                                 Longitude = -156.357567,
                                 Waterbody_name = "Ben Courtney Creek",
                                 Waterbody_type = "S"))

cik.meta <- bind_rows(cik.meta,
                      data.frame(Agency_ID = "Roadhouse Creek",
                                 SourceName = "CIK",
                                 Contact_person = "Sue Mauger",
                                 Latitude = 59.7573,
                                 Longitude = -154.8499,
                                 Waterbody_name = "Roadhouse Creek",
                                 Waterbody_type = "S"))

cik.meta <- cik.meta %>% 
  mutate(Agency_ID = case_when(grepl("Gibraltar", Agency_ID) ~ "Gibraltar River",
                            TRUE ~ Agency_ID))

cik.meta
```

Add SiteIDs to match with data that have CIK_ prefix.

```{r}

cik.meta <- cik.meta %>% 
  mutate(SiteID = paste0("cik_", Agency_ID))


saveRDS(cik.meta, "output/cik_metadata.rds")

```



# Data

Read in temp observation data from CIK folder

(Note checked error on logger out of water on silver salmon 2014, in another column so temp data will get an na there, all good.)

```{r}
cik.files <-list.files(path = paste0("data/CIK Bristol Bay village and lodge sites"),
               pattern = "*.xlsx", 
               full.names = T)

#remove metadata file
cik.files <- cik.files[!(grepl("CIK Bristol Bay village and lodge sites/CIK Bristol", cik.files))]


cik.data <- cik.files %>% 
  map_df(function(x) read_excel(x,skip = 1,col_names = F) %>% 
  mutate(file_name = gsub(".xlsx","",basename(x))))

cik.data <- cik.data %>% 
  mutate(sampleDate = as.Date(`...1`),
         sampleTime = as_hms(`...2`),
         Temperature = `...3`,
         Agency_ID = sub("\\_.*", "", file_name)) %>%  
  select(-`...1`, -`...2`, -`...3`, -`...4`, -`...5`) 

#fix site names. 
cik.data <- cik.data %>% 
  mutate(Agency_ID = case_when(grepl("Naknek", file_name) ~ "Big Creek_Naknek",
                            grepl("Roadhouse", file_name) ~ "Roadhouse Creek",
                            grepl("Panar", file_name) ~ "Panaruqak Creek",
                            TRUE ~ Agency_ID))


```

Add SiteIDs to match with data that have CIK_ prefix.

```{r}
cik.data <- cik.data %>% 
  mutate(SiteID = paste0("cik_", Agency_ID))

cik.data %>% 
  distinct(Agency_ID, SiteID) %>% 
  left_join(cik.meta)
```



Summary csv to add as attributes to leaflet map.

```{r}
cik.data %>% 
  group_by(SiteID) %>% 
  summarize(startYear = min(year(sampleDate)),
            endYear = max(year(sampleDate)),
            totYears = length(unique(year(sampleDate)))) %>% 
  saveRDS("output/cik_data_summ.rds")

```

```{r}
cik.data %>% 
  saveRDS("output/cik_data.rds")
```

Perform a quick visualization to see extent and form of data
```{r eval = FALSE}
cik.data %>%
  mutate(DT = parse_date_time(paste(sampleDate, sampleTime), "Y-m-d H:M:S")) %>%
  ggplot(aes(x = DT, y = Temperature)) +
  geom_point() +
  facet_wrap(~ SiteID) +
  ggtitle("Original Logger Data by Site and Year")

cik.data %>% distinct(SiteID)

```

# Save daily data and metadata

Save raw data for AKTEMP, which includes date and time. Save metadata file. Save daily data after screening for days with less than 90% of measurements.

```{r save raw data}
source("W:/Github/AKSSF/helper_functions.R")

cik.meta

save_metadata_files(cik.meta, "cik_bb")

cik.data

cik.data %>% 
  distinct(SiteID) %>% 
  left_join(cik.meta)

# save_aktemp_files(cik.data %>% mutate(UseData = 1), "cik_bb")

cik.daily <- temp_msmt_freq(cik.data %>% mutate(UseData = 1)) %>% daily_screen(.)

save_daily_files(cik.daily, "cik_bb")
```
